---
title: "cat( )"
subtitle: "Print labelled results with units to the console. See [original documentation](https://rdrr.io/r/base/cat.html) for reference."
---

## Syntax
Required arguments:
```{r, eval=FALSE}
cat(...)
```
```{r, eval=FALSE}
cat("Label:", value, "unit\n")
cat("Label:", signif(value, digits), "unit\n")
```
Advanced use — Required and optional arguments:
```{r, eval=FALSE}
cat(..., sep = " ", file = "", append = FALSE)
```

## Description

::: {.desc-item}
`cat(...)` concatenates and prints its arguments to the console with no formatting.
Unlike `print()`, it does not add quotes around character strings or display `[1]` index prefixes, making it ideal for producing clean, readable output.
:::

---

::: {.desc-item}
`cat("Label:", value, "unit\n")` is the standard pattern for reporting a named result with its unit.
The `\n` at the end moves the cursor to a new line.
:::

---

::: {.desc-item}
`cat("Label:", signif(value, digits), "unit\n")` rounds the value to the appropriate number of significant figures at the point of printing, while keeping the stored variable at full precision.
:::

::: {.callout-note}
## Coding convention — always label printed results
Never print a bare variable name or call `print()` on a calculated value.
Always use `cat()` with a descriptive label and the unit.
A number without context is meaningless in a pharmaceutical report.
:::

## Examples

### Print a single analytical result with units

```{webr-r}
# Full-precision intermediate — do NOT round here
std_intercept <- 0.003
std_slope     <- 0.01420   # AU per µg/mL
absorbance    <- 0.445     # measured absorbance of sample

C2_ug_ml <- (absorbance - std_intercept) / std_slope   # Beer–Lambert

# Only round at the point of printing
cat("Sample concentration:", signif(C2_ug_ml, 3), "µg/mL\n")
```

### Print multiple results from a content uniformity assay

```{webr-r}
library(dplyr)
library(tibble)

label_claim_mg    <- 75.0
ibu_per_mg_powder <- 0.43

capsules <- tibble(
  fill_mass_mg = c(174.2, 174.4, 174.7, 174.0, 174.5, 174.3)
) %>%
  mutate(
    ibu_mg    = fill_mass_mg * ibu_per_mg_powder,
    pct_label = ibu_mg / label_claim_mg * 100
  )

X_bar_pct <- mean(capsules$pct_label)
s_pct     <- sd(capsules$pct_label)

cat("Mean content (X̄):", signif(X_bar_pct, 4), "% of label claim\n")
cat("Std deviation (s):", signif(s_pct, 3),     "%\n")
cat("n:", nrow(capsules), "capsules\n")
```

### Print a pass/fail acceptance criterion

```{webr-r}
# EP 2.9.40 Stage 1: AV = |M - X̄| + k·s
M    <- 100.0   # reference value (% label claim)
k_s1 <- 2.4     # EP Table 2.9.40-1, Stage 1, n = 6
L1   <- 15.0    # acceptance limit Stage 1 (%)

X_bar_pct <- 98.4
s_pct     <- 1.8

AV_s1 <- abs(M - X_bar_pct) + k_s1 * s_pct

cat("Acceptance value (AV):", round(AV_s1, 2), "\n")
cat("Acceptance limit (L1):", L1, "\n")
cat("Stage 1 result:", ifelse(AV_s1 <= L1, "PASS", "FAIL"), "\n")
```

### `sep` — custom separator between items

```{webr-r}
conc_values_ug_ml <- c(2.5, 5.0, 10.0, 20.0)

# Default sep = " " (space between items)
cat("Standards (µg/mL):", conc_values_ug_ml, "\n")

# Comma-separated
cat("Standards (µg/mL):", conc_values_ug_ml, sep = ", ")
cat("\n")
```

## Input Arguments

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>...</code> — items to print <span class="arg-types">character | numeric | logical</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
One or more values to print, separated by commas.
Items are coerced to character before printing.
Common usage combines a text label, a computed value (wrapped in `signif()` or `round()`), and a unit string:

```r
cat("Mean fill mass:", signif(X_bar, 4), "mg\n")
```

Use `\n` (newline) as the last item so the next output starts on a fresh line.

<p class="data-types"><strong>Data Types:</strong> <code>character</code> | <code>numeric</code> | <code>logical</code> | <code>integer</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>sep</code> — separator between items <span class="arg-types">character</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
A character string inserted between consecutive items.
Default is a single space `" "`.
Set `sep = ""` to print items with no gap, or `sep = ", "` for comma-separated lists.

```r
cat("Result:", value, "mg\n")          # default: space between each item
cat("A", "B", "C", sep = "-")          # prints: A-B-C
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> · Default: <code>" "</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>file</code> — output destination <span class="arg-types">character | connection</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Where to send the output.
Default `""` prints to the console.
Pass a file path string (or an open connection) to write to a file instead.

```r
cat("Result:", signif(AV_s1, 4), "\n", file = "results.txt")
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> (file path) | <code>connection</code> · Default: <code>""</code> (console)</p>
</div>
</details>

<details class="arg-item">
<summary><code>append</code> — append to file instead of overwriting <span class="arg-types">logical</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Only relevant when `file` is a file path.
If `TRUE`, new output is appended to the end of the file.
If `FALSE` (default), the file is overwritten.

<p class="data-types"><strong>Data Types:</strong> <code>logical</code> · Default: <code>FALSE</code></p>
</div>
</details>
:::
