---
title: "cat( )"
subtitle: "Print labelled results with units to the console. cat() is a base R function."
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}
{{< include /_webr-options.qmd >}}

<div class="source-links">
<a class="source-link" href="https://rdrr.io/r/base/cat.html" target="_blank" rel="noopener">Original Documentation ↗</a>
</div>

## Required Library
```r
# cat() is part of base R — no package installation needed
```

## Syntax

```{r, eval=FALSE}
cat(...)
```
::: {.desc-item}
`cat(...)` concatenates and prints its arguments to the console with no formatting.
Unlike `print()`, it does not add quotes around character strings or display `[1]` index prefixes, making it ideal for producing clean, readable output.
:::

```{r, eval=FALSE}
cat("Label:", value, "unit\n")
```
::: {.desc-item}
`cat("Label:", value, "unit\n")` is the standard pattern for reporting a named result with its unit.
The `\n` at the end moves the cursor to a new line, creating line breaks. Useful for printing multiple results in a report.
:::

::: {.callout-note}
## Coding convention — always label printed results
Never print a bare variable name or call `print()` on a calculated value.
Always use `cat()` with a descriptive label and the unit.
A number without context is meaningless in a pharmaceutical report.
:::

## Argument Overview
<p class="arg-intro">Required arguments must be included when using a function while optional arguments can be included on demand.</p>

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>...</code> — items to print <span class="arg-types">character | numeric | logical</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
One or more values to print, separated by commas.
Items are coerced to character before printing.
Common usage combines a text label, a computed value (wrapped in `signif()` or `round()`), and a unit string:

```r
cat("Mean fill mass:", signif(X_bar, 4), "mg\n")
```

Use `\n` (newline) as the last item so the next output starts on a fresh line.

<p class="data-types"><strong>Data Types:</strong> <code>character</code> | <code>numeric</code> | <code>logical</code> | <code>integer</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>sep</code> — separator between items <span class="arg-types">character</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
A character string inserted between consecutive items.
Default is a single space `" "`.
Set `sep = ""` to print items with no gap, or `sep = ", "` for comma-separated lists.

```r
cat("Result:", value, "mg\n")          # default: space between each item
cat("A", "B", "C", sep = "-")          # prints: A-B-C
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> · Default: <code>" "</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>file</code> — output destination <span class="arg-types">character | connection</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Where to send the output.
Default `""` prints to the console.
Pass a file path string (or an open connection) to write to a file instead.

```r
cat("Result:", signif(AV_s1, 4), "\n", file = "results.txt")
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> (file path) | <code>connection</code> · Default: <code>""</code> (console)</p>
</div>
</details>

<details class="arg-item">
<summary><code>append</code> — append to file instead of overwriting <span class="arg-types">logical</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Only relevant when `file` is a file path.
If `TRUE`, new output is appended to the end of the file.
If `FALSE` (default), the file is overwritten.

<p class="data-types"><strong>Data Types:</strong> <code>logical</code> · Default: <code>FALSE</code></p>
</div>
</details>
:::

## Examples

<div class="example-section">
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="example-item">
<summary>Combine text and values</summary>
<div class="example-body">

```{webr}
value <- 12.345678

cat("Result: ", value, "mg")
```
Rounds the value to the appropriate number of significant figures at the point of printing, while keeping the stored variable at full precision.
</div>
</details>

<details class="example-item">
<summary>Print multiple results from a content uniformity assay</summary>
<div class="example-body">
`"\n"` creates a line break so each result is printed on its own line. 
```{webr}
library(tidyverse)

capsules <- tibble(
  sample_id = paste0("C", 1:12),
  measured_ibu_mg = c(75.2, 74.8, 75.5, 74.9, 75.1, 74.7, 75.3, 75.0, 74.6, 75.2, 75.1, 74.8)
)

X_bar_pct <- mean(capsules$measured_ibu_mg / 75 * 100)
s_pct <- sd(capsules$measured_ibu_mg / 75 * 100)

cat("Mean content (X̄):", X_bar_pct, "%\n")
cat("Std deviation (s):", s_pct, "%\n")
cat("n:", nrow(capsules), "capsules\n")
```
</div>
</details>


<details class="example-item">
<summary>Custom separator between items</summary>
<div class="example-body">

```{webr}
conc_values_ug_ml <- c(2.5, 5.0, 10.0, 20.0)

# Default sep = " " (space between items)
cat("Standards (µg/mL):", conc_values_ug_ml, "\n")

# Comma-separated
cat("Standards (µg/mL):", conc_values_ug_ml, sep = ", ")

# Enforce line break at the end if you want to print more results after this
cat("\n")
```
</div>
</details>

</div>
