---
title: "coef( )"
subtitle: "Extract model coefficients. coef() is a generic function in base R (stats package)."
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}
{{< include /_webr-options.qmd >}}

<div class="source-links">
<a class="source-link" href="https://stat.ethz.ch/R-manual/R-devel/library/stats/html/coef.html" target="_blank" rel="noopener">Original Documentation ↗</a>
<a class="source-link" href="https://www.rdocumentation.org/packages/stats/topics/coef" target="_blank" rel="noopener">R Documentation ↗</a>
</div>

## Required Library
```r
# coef() is in base R (stats), loaded by default — no package installation needed
```

## Syntax

```{r, eval=FALSE}
coef(object)
```
::: {.desc-item}
`coef(object)` extracts estimated coefficients from a fitted model object.
For a simple linear model, this usually returns intercept and slope.
:::

```{r, eval=FALSE}
coefs <- coef(model)

intercept <- coefs[[1]]
slope <- coefs[[2]]
```
::: {.desc-item}
Use indexing when you need individual coefficient values for calculations, plotting, or formatted output.
:::

::: {.callout-tip collapse="true"}
## `coef()` vs `summary(model)$coefficients`
Use `coef(model)` when you only need the coefficient estimates.
Use `summary(model)$coefficients` when you also need standard errors, t-values, and p-values.
:::

## Argument Overview
<p class="arg-intro">Required arguments must be included when using a function while optional arguments can be included on demand.</p>

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>object</code> — fitted model object <span class="arg-types">model object</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
The fitted object from which coefficients should be extracted.
Commonly produced by `lm()`, but `coef()` also works for many other model classes.

```r
model <- lm(absorbance ~ conc_ug_ml, data = calibration)
coef(model)
```

<p class="data-types"><strong>Data Types:</strong> model object (for example <code>lm</code>, <code>glm</code>, <code>nls</code>)</p>
</div>
</details>

<details class="arg-item">
<summary><code>...</code> — method-specific options <span class="arg-types">various</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Additional arguments passed to class-specific methods.
For many everyday `lm()` uses, you do not need to set any extra arguments.


<p class="data-types"><strong>Data Types:</strong> depends on model class</p>
</div>
</details>
:::

## Examples

<div class="example-section">
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="example-item">
<summary>Extract intercept and slope from an lm() model</summary>
<div class="example-body">

```{webr}
library(tidyverse)

calibration <- tibble(
  conc_ug_ml = c(0, 2.5, 5.0, 10.0, 20.0, 40.0),
  absorbance = c(0.003, 0.057, 0.110, 0.226, 0.441, 0.889)
)

model <- lm(absorbance ~ conc_ug_ml, data = calibration)

coefs <- coef(model)
intercept <- coefs[[1]]
slope <- coefs[[2]]

cat("Intercept:", signif(intercept, 4), "\n")
cat("Slope:    ", signif(slope, 4), "\n")
```
</div>
</details>

<details class="example-item">
<summary>Access coefficients by name</summary>
<div class="example-body">

```{webr}
library(tidyverse)

calibration <- tibble(
  conc_ug_ml = c(0, 2.5, 5.0, 10.0, 20.0, 40.0),
  absorbance = c(0.003, 0.057, 0.110, 0.226, 0.441, 0.889)
)

model <- lm(absorbance ~ conc_ug_ml, data = calibration)
coefs <- coef(model)

coefs

intercept <- coefs["(Intercept)"]
slope <- coefs["conc_ug_ml"]

cat("Intercept (named):", signif(intercept, 4), "\n")
cat("Slope (named):    ", signif(slope, 4), "\n")
```
</div>
</details>

<details class="example-item">
<summary>Use coef() values to draw a regression line in ggplot</summary>
<div class="example-body">

```{webr}
library(tidyverse)

calibration <- tibble(
  conc_ug_ml = c(0, 2.5, 5.0, 10.0, 20.0, 40.0),
  absorbance = c(0.003, 0.057, 0.110, 0.226, 0.441, 0.889)
)

model <- lm(absorbance ~ conc_ug_ml, data = calibration)
coefs <- coef(model)

ggplot(calibration, aes(x = conc_ug_ml, y = absorbance)) +
  geom_point(size = 2) +
  geom_abline(
    intercept = coefs[[1]],
    slope = coefs[[2]],
    colour = "#3C5E3E",
    linewidth = 1
  ) +
  labs(
    title = "Regression Line from coef()",
    x = "Concentration (ug/mL)",
    y = "Absorbance"
  ) +
  theme_classic()
```
</div>
</details>

<details class="example-item">
<summary>Compare coef() with summary() coefficient table</summary>
<div class="example-body">

```{webr}
library(tidyverse)

calibration <- tibble(
  conc_ug_ml = c(0, 2.5, 5.0, 10.0, 20.0, 40.0),
  absorbance = c(0.003, 0.057, 0.110, 0.226, 0.441, 0.889)
)

model <- lm(absorbance ~ conc_ug_ml, data = calibration)

cat("coef(model):\n")
print(coef(model))

cat("\nsummary(model)$coefficients:\n")
print(summary(model)$coefficients)
```
</div>
</details>

</div>