---
title: "ggplot( )"
subtitle: "Create a plot by mapping data variables to visual aesthetics. ggplot() is from the ggplot2 package, which is part of the tidyverse."
---

<div class="source-links">
<a class="source-link" href="https://ggplot2.tidyverse.org/reference/ggplot.html" target="_blank" rel="noopener">Original Documentation ↗</a>
<a class="source-link" href="https://rstudio.github.io/cheatsheets/html/data-visualization.html" target="_blank" rel="noopener">ggplot2 Cheatsheet ↗</a>
</div>

## Required Library
```r
install.packages("tidyverse")
library(tidyverse)
```

## Syntax

```{r, eval=FALSE}
# Minimal pattern
ggplot(data = df, aes(x = x_var, y = y_var)) +
  geom_point()
```
::: {.desc-item}
`ggplot(data = ..., aes(...))` initializes a plot object.
You then add one or more `geom_*()` layers with `+` to draw points, lines, bars, and other marks.
:::

```{r, eval=FALSE}
# Pipe style
df %>%
  ggplot(aes(x = x_var, y = y_var, color = group_var)) +
  geom_line()
```
::: {.desc-item}
You can pass a data frame to `ggplot()` through a pipe.
Inside `aes()`, map variables to aesthetics such as position (`x`, `y`), color, shape, or size.
:::

::: {.callout-tip collapse="true"}
## Why `+` and not `%>%` between layers?
`ggplot()` builds plots by adding layers to a plot object, so each `+` appends a new component.
Use `%>%` (or `|>`) to prepare data before plotting, then use `+` for plot layers.
:::

## Argument Overview
<p class="arg-intro">Required arguments must be included when using a function while optional arguments can be included on demand.</p>

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>data</code> — data frame used for plotting <span class="arg-types">data frame | tibble</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
The data frame that contains variables used in the plot.
If omitted, data can be supplied in a later layer (for example inside `geom_point(data = ...)`).

```r
ggplot(data = capsules, aes(x = time_min, y = pct_release))
```

<p class="data-types"><strong>Data Types:</strong> <code>data.frame</code> | <code>tibble</code> · Default: <code>NULL</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>mapping</code> — aesthetic mappings made with <code>aes()</code> <span class="arg-types">uneval mapping</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Defines how variables map to visual properties like `x`, `y`, `color`, `shape`, `linetype`, or `size`.
Aesthetic mappings can be set globally in `ggplot()` and/or locally inside each layer.

```r
ggplot(capsules, aes(x = batch, y = pct_label, color = batch))
```

<p class="data-types"><strong>Data Types:</strong> <code>aes()</code> mapping object · Default: <code>aes()</code> (empty mapping)</p>
</div>
</details>

<details class="arg-item">
<summary><code>...</code> — additional parameters <span class="arg-types">varies</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Additional options forwarded internally. Most day-to-day plot control is usually done in layers and helper functions (for example `geom_*()`, `labs()`, `theme_*()`, and `scale_*()`).

```r
ggplot(capsules, aes(time_min, pct_release)) +
  geom_line(linewidth = 1)
```

<p class="data-types"><strong>Data Types:</strong> varies by use case</p>
</div>
</details>
:::

## Examples

<div class="example-section">
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="example-item">
<summary>Scatter plot of tablet hardness vs. disintegration time</summary>
<div class="example-body">

```{webr-r}
library(tidyverse)

tablets <- tibble(
  tablet_id = 1:8,
  hardness_n = c(62, 59, 67, 64, 61, 66, 63, 68),
  disintegration_min = c(8.9, 9.4, 7.8, 8.2, 9.0, 7.6, 8.5, 7.2)
)

ggplot(tablets, aes(x = hardness_n, y = disintegration_min)) +
  geom_point(size = 2.8, alpha = 0.9) +
  labs(
    title = "Tablet hardness vs disintegration time",
    x = "Hardness (N)",
    y = "Disintegration time (min)"
  )
```
</div>
</details>

<details class="example-item">
<summary>Line plot of dissolution profile over time</summary>
<div class="example-body">

```{webr-r}
library(tidyverse)

dissolution <- tibble(
  time_min = c(5, 10, 15, 20, 30, 45, 60),
  pct_release = c(22, 39, 55, 66, 79, 90, 96)
)

ggplot(dissolution, aes(x = time_min, y = pct_release)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title = "Dissolution profile",
    x = "Time (min)",
    y = "% released"
  )
```
</div>
</details>

<details class="example-item">
<summary>Grouped boxplot by batch</summary>
<div class="example-body">

```{webr-r}
library(tidyverse)

content <- tibble(
  batch = rep(c("A", "B", "C"), each = 8),
  pct_label = c(
    99.2, 101.3, 98.7, 100.1, 102.0, 99.5, 100.8, 98.9,
    96.1, 97.4, 98.2, 95.8, 97.0, 96.9, 98.0, 97.2,
    101.5, 103.2, 102.1, 100.9, 101.8, 103.0, 102.7, 101.2
  )
)

ggplot(content, aes(x = batch, y = pct_label, fill = batch)) +
  geom_boxplot(alpha = 0.8, width = 0.65, outlier.shape = 16, outlier.size = 2) +
  guides(fill = "none") +
  labs(
    title = "Content uniformity by batch",
    x = "Batch",
    y = "% of label claim"
  )
```
</div>
</details>

</div>
