---
title: "lm( )"
subtitle: "Fit linear regression models. lm() is part of base R (stats package)."
---

<div class="source-links">
<a class="source-link" href="https://stat.ethz.ch/R-manual/R-devel/library/stats/html/lm.html" target="_blank" rel="noopener">Original Documentation ↗</a>
<a class="source-link" href="https://www.rdocumentation.org/packages/stats/topics/lm" target="_blank" rel="noopener">R Documentation ↗</a>
</div>

## Required Library
```r
# lm() is in base R (stats), loaded by default — no package installation needed
```

## Syntax

```{r, eval=FALSE}
lm(y ~ x, data = df)

# Pipe style (common in tidyverse workflows)
df %>% lm(y ~ x, data = .)

```
::: {.desc-item}
`lm(y ~ x, data = df)` fits a straight-line model in the form `y = ax + b`, where `a` is slope and `b` is intercept.
In formula syntax, the left side is the response and the right side is the predictor(s).
:::

```{r, eval=FALSE}
lm(y ~ 0 + x, data = df)
lm(y ~ x - 1, data = df)
```
::: {.desc-item}
The above variants fit a zero-intercept model (forced through the origin).
Use this only when the scientific model justifies it.
:::

::: {.callout-tip collapse="true"}
## Use helper functions on objects created by `lm()`?
`lm()` returns a model object. Use helper functions to extract what you need:

- `coef(model)` for intercept and slope
- `summary(model)$r.squared` for goodness of fit
- `predict(model, newdata = ...)` for predictions
:::

## Argument Overview
<p class="arg-intro">Required arguments must be included when using a function while optional arguments can be included on demand.</p>

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>formula</code> — model formula <span class="arg-types">formula</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
Specifies the model structure, usually `y ~ x` for simple linear regression.
Left side is the response variable; right side is the predictor.

```r
lm(absorbance ~ conc_ug_ml, data = calibration)
lm(absorbance ~ 0 + conc_ug_ml, data = calibration)
```

<p class="data-types"><strong>Data Types:</strong> <code>formula</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>data</code> — data frame containing variables <span class="arg-types">data frame | tibble</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Data source used to evaluate variables in the formula.

This argument can be omitted, but then the variables in the formula (for example `y` and `x`) must already exist as objects in your environment.

```r
lm(absorbance ~ conc_ug_ml, data = calibration)

# Also valid if absorbance and conc_ug_ml already exist as standalone vectors:
lm(absorbance ~ conc_ug_ml)
```

<p class="data-types"><strong>Data Types:</strong> <code>data.frame</code> | <code>tibble</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>subset</code> — rows to include in fit <span class="arg-types">logical | integer</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Fits the model on a selected subset of rows.

```r
lm(absorbance ~ conc_ug_ml, data = calibration, subset = conc_ug_ml <= 20)
```

<p class="data-types"><strong>Data Types:</strong> logical or integer row index · Default: all rows</p>
</div>
</details>

<details class="arg-item">
<summary><code>weights</code> — observation weights <span class="arg-types">numeric</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Relative weights for weighted least squares (WLS).

```r
lm(absorbance ~ conc_ug_ml, data = calibration, weights = wt)
```

<p class="data-types"><strong>Data Types:</strong> numeric vector (same length as observations) · Default: <code>NULL</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>na.action</code> — missing-value handling <span class="arg-types">function</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Controls how missing values are handled during model fitting.

```r
lm(absorbance ~ conc_ug_ml, data = calibration, na.action = na.exclude)
```

<p class="data-types"><strong>Data Types:</strong> function (for example <code>na.omit</code>, <code>na.exclude</code>) · Default: session option</p>
</div>
</details>
:::

## Examples

<div class="example-section">
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="example-item">
<summary>Fit a calibration line and extract slope/intercept</summary>
<div class="example-body">

```{webr-r}
library(tidyverse)

calibration <- tibble(
  conc_ug_ml = c(0, 2.5, 5.0, 10.0, 20.0, 40.0),
  absorbance = c(0.003, 0.057, 0.110, 0.226, 0.441, 0.889)
)

model <- lm(absorbance ~ conc_ug_ml, data = calibration)

coefs <- coef(model)
intercept <- coefs[[1]]
slope <- coefs[[2]]

cat("Intercept:", signif(intercept, 4), "\n")
cat("Slope:    ", signif(slope, 4), "\n")
```
</div>
</details>

<details class="example-item">
<summary>Extract R-squared and inspect model summary</summary>
<div class="example-body">

```{webr-r}
library(tidyverse)

calibration <- tibble(
  conc_ug_ml = c(0, 2.5, 5.0, 10.0, 20.0, 40.0),
  absorbance = c(0.003, 0.057, 0.110, 0.226, 0.441, 0.889)
)

model <- lm(absorbance ~ conc_ug_ml, data = calibration)
model_summary <- summary(model)

r2 <- model_summary$r.squared
cat("R-squared:", signif(r2, 4), "\n")

# Optional: full printed report
model_summary
```
</div>
</details>

<details class="example-item">
<summary>Predict absorbance at new concentration values</summary>
<div class="example-body">

```{webr-r}
library(tidyverse)

calibration <- tibble(
  conc_ug_ml = c(0, 2.5, 5.0, 10.0, 20.0, 40.0),
  absorbance = c(0.003, 0.057, 0.110, 0.226, 0.441, 0.889)
)

model <- lm(absorbance ~ conc_ug_ml, data = calibration)

new_conc <- tibble(conc_ug_ml = c(7.5, 15, 30))
pred_abs <- predict(model, newdata = new_conc)

# Combine predictions with new concentration values
bind_cols(new_conc, tibble(pred_absorbance = pred_abs))
```
</div>
</details>

<details class="example-item">
<summary>Add the fitted lm() line to a ggplot with geom_abline()</summary>
<div class="example-body">

```{webr-r}
library(tidyverse)

calibration <- tibble(
  conc_ug_ml = c(0, 2.5, 5.0, 10.0, 20.0, 40.0),
  absorbance = c(0.003, 0.057, 0.110, 0.226, 0.441, 0.889)
)

model <- lm(absorbance ~ conc_ug_ml, data = calibration)

coefs <- coef(model)
intercept <- coefs[[1]]
slope <- coefs[[2]]

ggplot(calibration, aes(x = conc_ug_ml, y = absorbance)) +
  geom_point(size = 2) +
  geom_abline(intercept = intercept, slope = slope, colour = "#3C5E3E", linewidth = 1) +
  labs(
    title = "Calibration Data with lm() Fit",
    x = "Concentration (ug/mL)",
    y = "Absorbance"
  ) +
  theme_classic()
```
</div>
</details>

<details class="example-item">
<summary>Compare free-intercept and zero-intercept models</summary>
<div class="example-body">

```{webr-r}
library(tidyverse)

calibration <- tibble(
  conc_ug_ml = c(0, 2.5, 5.0, 10.0, 20.0, 40.0),
  absorbance = c(0.003, 0.057, 0.110, 0.226, 0.441, 0.889)
)

model_free <- lm(absorbance ~ conc_ug_ml, data = calibration)
model_zero <- lm(absorbance ~ 0 + conc_ug_ml, data = calibration)

cat("Free intercept model:\n")
print(coef(model_free))

cat("\nZero-intercept model:\n")
print(coef(model_zero))
```
</div>
</details>

</div>