---
title: "mutate( )"
subtitle: "Add or transform columns in a data frame. See [original documentation](https://dplyr.tidyverse.org/reference/mutate.html) for reference."
---

## Syntax
Required arguments:
```{r, eval=FALSE}
df %>% mutate(new_col = expression)
```
```{r, eval=FALSE}
df %>% mutate(new_col = expression, another_col = expression)
```
Advanced use — Required and optional arguments:
```{r, eval=FALSE}
df %>% mutate(..., .keep = "all", .before = NULL, .after = NULL)
```

## Description

::: {.desc-item}
`df %>% mutate(new_col = expression)` adds a new column to a data frame.
The expression on the right can reference any existing column by name.
:::

---

::: {.desc-item}
`df %>% mutate(col1 = ..., col2 = ...)` adds multiple columns in a single call.
Later expressions can reference columns created earlier in the same `mutate()`.
:::

---

::: {.desc-item}
`df %>% mutate(..., .keep = "all")` controls which columns are kept in the result.
The default `"all"` retains every column, including unmodified ones.
:::

::: {.callout-note}
## Coding convention — never round inside `mutate()`
Store full-precision values in all columns.
Apply `signif()` or `round()` only inside `cat()` when printing a final result.
Rounding a column that feeds into later calculations accumulates error.
:::

## Examples

### Capsule content uniformity — computing fill mass and % label claim

```{webr-r}
library(dplyr)
library(tibble)

# Constants — update these with your values
label_claim_mg    <- 75.0   # mg — labelled dose per capsule
ibu_per_mg_powder <- 0.43   # mg ibuprofen per mg powder (from UV assay)

# Raw weighing data
capsules <- tibble(
  capsule_id     = 1:6,
  mass_empty_mg  = c(91.2, 88.7, 90.1, 89.4, 92.0, 87.8),
  mass_filled_mg = c(265.4, 263.1, 264.8, 263.9, 266.1, 262.7)
)

capsules <- capsules %>%
  mutate(
    fill_mass_mg = mass_filled_mg - mass_empty_mg,         # net fill weight
    ibu_mg       = fill_mass_mg * ibu_per_mg_powder,       # ibuprofen per capsule
    pct_label    = ibu_mg / label_claim_mg * 100           # % of label claim
  )

capsules
```

### Flagging out-of-specification results

Use `mutate()` with `ifelse()` to add a pass/fail column based on a criterion.

```{webr-r}
library(dplyr)

# EP 2.9.40 Stage 1: each unit must be within ±25 % of label claim (M ± 25)
# Here we flag units outside 85–115 % as OOS (simplified uniform-content criterion)
capsules <- capsules %>%
  mutate(
    oos = ifelse(pct_label < 85 | pct_label > 115, "OOS", "Pass")
  )

capsules[, c("capsule_id", "pct_label", "oos")]
```

### Concentration units conversion

```{webr-r}
library(dplyr)
library(tibble)

# Absorbance readings from a dissolution sample series
samples <- tibble(
  time_min    = c(15, 30, 45, 60, 90, 120),
  absorbance  = c(0.142, 0.287, 0.401, 0.489, 0.531, 0.558)
)

# Standard curve: C (µg/mL) = (A - intercept) / slope
std_intercept <- 0.003    # y-intercept of calibration curve
std_slope     <- 0.01420  # slope (AU per µg/mL)
vol_flask_mL  <- 50.0     # dissolution vessel volume
dose_mg       <- 75.0     # labelled dose

samples <- samples %>%
  mutate(
    conc_ug_ml  = (absorbance - std_intercept) / std_slope,   # Beer–Lambert inversion
    conc_mg     = conc_ug_ml * vol_flask_mL / 1000,           # µg/mL → mg in vessel
    pct_label   = conc_mg / dose_mg * 100                     # % label claim dissolved
  )

samples
```

## Input Arguments

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>.data</code> <span class="arg-types">data frame | tibble</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
The data frame to modify. In a pipe (`%>%` or `%>%`), this is passed automatically from the left-hand side — you do not write it explicitly.

```r
# .data is passed automatically by the pipe:
capsules %>% mutate(fill_mass_mg = mass_filled_mg - mass_empty_mg)
```

<p class="data-types"><strong>Data Types:</strong> <code>data.frame</code> | <code>tibble</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>...</code> — name–expression pairs <span class="arg-types">name = expr</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
One or more expressions of the form `new_column_name = expression`.
The right-hand side can reference any column in `.data` by name, and can also reference columns created earlier in the same `mutate()` call.
Separate multiple expressions with commas.

```r
mutate(
  fill_mass_mg = mass_filled_mg - mass_empty_mg,   # uses two existing columns
  ibu_mg       = fill_mass_mg * ibu_per_mg_powder  # references the column just created above
)
```

<p class="data-types"><strong>Data Types:</strong> any expression that returns a vector the same length as the data frame, or a single value (recycled to all rows)</p>
</div>
</details>

<details class="arg-item">
<summary><code>.keep</code> <span class="arg-types">character</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Controls which columns from `.data` appear in the output alongside the new/modified columns.

| Value | Behaviour |
|---|---|
| `"all"` | Keep all original columns **(default)** |
| `"used"` | Keep only columns referenced in the expressions |
| `"unused"` | Keep only columns *not* referenced |
| `"none"` | Drop all original columns (equivalent to `transmute()`) |

<p class="data-types"><strong>Data Types:</strong> <code>character</code> — one of <code>"all"</code>, <code>"used"</code>, <code>"unused"</code>, <code>"none"</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>.before</code> / <code>.after</code> <span class="arg-types">column name | integer</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Controls where the new columns are inserted.
By default new columns are appended on the right.

```r
capsules %>% mutate(fill_mass_mg = mass_filled_mg - mass_empty_mg, .after = mass_filled_mg)
```

<p class="data-types"><strong>Data Types:</strong> unquoted column name or integer position</p>
</div>
</details>
:::
