---
title: "mutate( )"
subtitle: "Add or transform columns in a data frame. mutate() is from the dplyr package, which is part of the tidyverse."
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}
{{< include /_webr-options.qmd >}}

<div class="source-links">
<a class="source-link" href="https://dplyr.tidyverse.org/reference/mutate.html" target="_blank" rel="noopener">Original Documentation ↗</a>
<a class="source-link" href="https://rstudio.github.io/cheatsheets/html/data-transformation.html" target="_blank" rel="noopener">dplyr Cheatsheet ↗</a>
</div>

## Required Library
```r
install.packages("tidyverse")
library(tidyverse)
```

## Syntax

```{r, eval=FALSE}
# Pipe style
df %>% mutate(new_col = expression)

# Non-pipe style
mutate(df, new_col = expression)
```
::: {.desc-item}
`df %>% mutate(new_col = expression)` and `mutate(df, new_col = expression)` add a new column to a data frame.
The expression on the right can reference any existing column by name.
:::

```{r, eval=FALSE}
# Pipe style
df %>% mutate(new_col = expression, another_col = expression)

# Non-pipe style
mutate(df, new_col = expression, another_col = expression)
```
::: {.desc-item}
`df %>% mutate(col1 = ..., col2 = ...)` and `mutate(df, col1 = ..., col2 = ...)` add multiple columns in a single call.
Later expressions can reference columns created earlier in the same `mutate()`.
:::

::: {.callout-tip collapse="true"}
## Why Are There Two Syntax Styles?
Both versions do the same thing.
With the pipe `%>%`, the data frame on the left is passed automatically as the first argument to `mutate()`.
Without the pipe, you write that first argument explicitly as `mutate(df, ...)`.
Use whichever style is clearer for your workflow.
:::



::: {.callout-note collapse="true"}
## Coding convention — never round inside `mutate()`
Store full-precision values in all columns.
Apply `signif()` or `round()` only inside `cat()` when printing a final result.
Rounding a column that feeds into later calculations accumulates error.
:::

## Argument Overview
<p class="arg-intro">Required arguments must be included when using a function while optional arguments can be included on demand.</p>

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>.data</code> <span class="arg-types">data frame | tibble</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
The data frame to modify. In a pipe (`%>%` or `%>%`), this is passed automatically from the left-hand side — you do not write it explicitly.

```r
# .data is passed automatically by the pipe:
capsules %>% mutate(fill_mass_mg = mass_filled_mg - mass_empty_mg)
```

<p class="data-types"><strong>Data Types:</strong> <code>data.frame</code> | <code>tibble</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>...</code> — name–expression pairs <span class="arg-types">name = expr</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
One or more expressions of the form `new_column_name = expression`.
The right-hand side can reference any column in `.data` by name, and can also reference columns created earlier in the same `mutate()` call.
Separate multiple expressions with commas.

```r
mutate(
  fill_mass_mg = mass_filled_mg - mass_empty_mg,   # uses two existing columns
  ibu_mg       = fill_mass_mg * ibu_per_mg_powder  # references the column just created above
)
```

<p class="data-types"><strong>Data Types:</strong> any expression that returns a vector the same length as the data frame, or a single value (recycled to all rows)</p>
</div>
</details>
:::

## Examples

<div class="example-section">
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="example-item">
<summary>Add a new column</summary>
<div class="example-body">

## Original data
```{webr}
library(tidyverse)

tablet_weights <- read_csv("data/tablet_weights.csv")

head(tablet_weights)
```

## Add a new column

```{webr}
library(tidyverse)

tablet_weights <- read_csv("data/tablet_weights.csv")

tablet_weights %>%
  mutate(weight_g = weight_mg / 1000)

```
</div>
</details>

<details class="example-item">
<summary>Add several columns at once</summary>
<div class="example-body">

```{webr}
library(tidyverse)

tablet_weights <- read_csv("data/tablet_weights.csv")

tablet_weights %>%
  mutate(
    weight_g   = weight_mg / 1000,
    dev_mg     = weight_mg - 500,
    pct_target = weight_mg / 500 * 100
  )

```
</div>
</details>

<details class="example-item">
<summary>Use a column you just created</summary>
<div class="example-body">

You can use a column you just created because mutate reads left to right.

```{webr}
library(tidyverse)

tablet_weights <- read_csv("data/tablet_weights.csv")

tablet_weights %>%
  mutate(
    dev_mg     = weight_mg - 500,
    abs_dev_mg = abs(dev_mg)
  )

```
</div>
</details>

<details class="example-item">
<summary>A conditional column with if_else()</summary>
<div class="example-body">

```{webr}
library(tidyverse)

tablet_weights <- read_csv("data/tablet_weights.csv")

tablet_weights %>%
  mutate(within_spec = if_else(abs(weight_mg - 500) <= 15, "pass", "fail"))

```
</div>
</details>

</div>
