---
title: "read_excel( )"
subtitle: "Import Excel worksheets as a tibble. read_excel(), read_xls(), and read_xlsx() are from the readxl package."
---

<div class="source-links">
<a class="source-link" href="https://readxl.tidyverse.org/reference/read_excel.html" target="_blank" rel="noopener">Original Documentation ↗</a>
<a class="source-link" href="https://rstudio.github.io/cheatsheets/html/data-import.html" target="_blank" rel="noopener">readxl Cheatsheet ↗</a>
</div>

## Required Library
read_excel() is part of the tidyverse ecosystem, but it is not included in the core tidyverse package. Install and load the readxl package before using read_excel():
```r
install.packages("readxl")
library(readxl)
```

## Syntax

```{r, eval=FALSE}
read_excel(path)
```
::: {.desc-item}
`read_excel("file_name.xlsx")` reads an Excel file (`.xls` or `.xlsx`) and returns a tibble data frame. By default, the first sheet of the Excel file is read, but you can specify a different sheet by name or position (see arguments below).

![](../assets/images/excel_to_tibble.svg)
:::

## Examples

<div class="example-section">
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>


<details class="example-item">
<summary>read_xls() and read_xlsx()</summary>
<div class="example-body">
```{r, eval=FALSE}
read_xls(path)
```
::: {.desc-item}
`read_xls("file_name.xls")` reads legacy Excel 97-2003 `.xls` files.
Use this when the file extension is `.xls` or the workbook was saved in the older binary format.
:::

<hr>

```{r, eval=FALSE}
read_xlsx(path)
```
::: {.desc-item}
`read_xlsx("file_name.xlsx")` reads modern Excel `.xlsx` files.
Use this when the workbook is in Office Open XML format (`.xlsx`).
:::

::: {.callout-tip collapse="true"}
## Which function should I use?
If you are not sure, use `read_excel()`.
It auto-detects whether your file is `.xls` or `.xlsx` from the extension.
Use `read_xls()` or `read_xlsx()` only when you want to force one file type.
:::

</div>
</details>

<details class="example-item">
<summary>Read the first worksheet from an Excel file</summary>
<div class="example-body">

```{webr}
library(readxl)

df <- read_excel("data/lab_results.xlsx")
head(df)
```
</div>
</details>

<details class="example-item">
<summary>Read a specific worksheet</summary>
<div class="example-body">

## Select a worksheet by name.

```{webr}
library(readxl)

df <- read_excel("data/lab_results.xlsx", sheet = "Dissolution")
head(df)
```

## Select a worksheet by position (e.g., second sheet).

```{webr}
library(readxl)

df <- read_excel("data/lab_results.xlsx", sheet = 2)
head(df)
```

</div>
</details>

<details class="example-item">
<summary>Import only a selected cell range</summary>
<div class="example-body">

```{webr}
library(readxl)

df <- read_excel("data/lab_results.xlsx", range = "A1:E25")
head(df)
```

## You can also combine sheet and range:

```{webr}
library(readxl)

df <- read_excel("data/lab_results.xlsx", sheet = "Dissolution", range = "A1:E25")
head(df)
```

</div>
</details>


<details class="example-item">
<summary>Skip rows</summary>
<div class="example-body">

This dataset has 3 title rows above the header:

```{webr}
library(readxl)
df <- read_excel("data/lab_results_titled.xlsx")
head(df)
```

Use `skip = 3` to ignore them:

```{webr}
library(readxl)
df <- read_excel("data/lab_results_titled.xlsx", skip = 3)
head(df)
```


</div>
</details>

<details class="example-item">
<summary>Handle custom missing value codes during import</summary>
<div class="example-body">

The sheet `Weights (missing)` contains missing values that were entered in the spreadsheet as `"N/A"`, `"-"`, and `"ND"`. They can be converted to `NA` during import by specifying them in the `na` argument:

```{webr}
library(readxl)

df <- read_excel(
  "data/lab_results.xlsx",
  sheet = "Weights (missing)",
  na = c("", "N/A", "-", "ND")
)

head(df)
```
</div>
</details>

</div>

## Argument Overview
<p class="arg-intro">Required arguments must be included when using a function while optional arguments can be included on demand.</p>

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>path</code> — path to the Excel file <span class="arg-types">character</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
A string giving the path to the Excel workbook.
Use forward slashes (`/`) in file paths, even on Windows.

```r
read_excel("data/results.xlsx")
read_excel("C:/Users/Morten/Documents/results.xlsx")
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> (file path)</p>
</div>
</details>

<details class="arg-item">
<summary><code>sheet</code> — worksheet to import <span class="arg-types">character | integer</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Choose which worksheet to read by name or by position.
Default is the first sheet.

```r
read_excel("results.xlsx", sheet = "Dissolution")
read_excel("results.xlsx", sheet = 2)
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> | <code>integer</code> · Default: first sheet</p>
</div>
</details>

<details class="arg-item">
<summary><code>range</code> — cell range to read <span class="arg-types">character</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Read a specific rectangular cell range using Excel notation.
Useful when the worksheet contains extra notes or multiple tables.

```r
read_excel("results.xlsx", range = "A1:D20")
read_excel("results.xlsx", range = "B3:F15")
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> (Excel-style range)</p>
</div>
</details>

<details class="arg-item">
<summary><code>col_names</code> — use first row as column names <span class="arg-types">logical | character</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
If `TRUE` (default), first row becomes column names.
Set to `FALSE` to auto-generate names, or pass a character vector to define names manually.

```r
read_excel("results.xlsx", col_names = FALSE)
read_excel("results.xlsx", col_names = c("tablet", "time_min", "absorbance"))
```

<p class="data-types"><strong>Data Types:</strong> <code>logical</code> | <code>character</code> vector · Default: <code>TRUE</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>na</code> — strings treated as missing values <span class="arg-types">character vector</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Defines which strings should be converted to `NA` while importing.
Useful when data entry used placeholders like `"N/A"` or `"-"`.

```r
read_excel("results.xlsx", na = c("", "N/A", "-", "Missing"))
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> vector · Default: <code>""</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>skip</code> — number of rows to skip <span class="arg-types">numeric</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Skips rows from the top before reading data.
Useful for spreadsheets with titles or metadata above the header row.

```r
read_excel("results.xlsx", skip = 2)
```

<p class="data-types"><strong>Data Types:</strong> <code>numeric</code> · Default: <code>0</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>n_max</code> — maximum rows to read <span class="arg-types">numeric</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Limits how many rows are imported.
Useful for previewing large workbooks.

```r
read_excel("results.xlsx", n_max = 100)
```

<p class="data-types"><strong>Data Types:</strong> <code>numeric</code> · Default: <code>Inf</code></p>
</div>
</details>

:::
