---
title: "read_csv( )"
subtitle: "Import a CSV file as a [tibble](tibble.qmd). readcsv() and readcsv2() are both specific implementations of the more general read_delim() function from the readr package."
---

<a class="source-link" href="https://readr.tidyverse.org/reference/read_delim.html" target="_blank" rel="noopener">Original Documentation ↗</a>

## Syntax

### Required Arguments

```{r, eval=FALSE}
read_csv(file)
```

```{r, eval=FALSE}
read_csv2(file)
```

### Description

::: {.desc-item}
`read_csv("file_name.csv")` reads a comma-separated values (.csv) file and returns a [tibble](tibble.qmd). Use if your file uses commas as column separators and dots for decimals.
:::
::: {.desc-item}
`read_csv2("file_name.csv")` reads a .csv file that uses semicolons as column separators and commas for decimals.
:::

::: {.callout-tip collapse="true"}
## Why 2 different functions?
English standard vs. danish standard. To find out how your `.csv`file is formatted, open it with a text editor on your computer or use the following code to print the first rows with R: `read_lines("file.csv", n_max = 3)`

```{webr-r}
read_lines("https://raw.githubusercontent.com/CPDSE-EDUX/SDU-CodeClub/main/debugging/dissolution_data.csv", n_max = 3)
```
:::

### Required and Optional Arguments

```{r, eval=FALSE}
read_csv(
  file,
  col_names = TRUE,
  col_types = NULL,
  col_select = NULL,
  id = NULL,
  locale = default_locale(),
  na = c("", "NA"),
  quoted_na = TRUE,
  quote = "\"",
  comment = "",
  trim_ws = TRUE,
  skip = 0,
  n_max = Inf,
  guess_max = min(1000, n_max),
  name_repair = "unique",
  num_threads = readr_threads(),
  progress = show_progress(),
  show_col_types = should_show_types(),
  skip_empty_rows = TRUE,
  lazy = should_read_lazy()
)
```

## Argument Overview

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item" open>
<summary><code>file</code> — path to the CSV file <span class="arg-types">character | connection</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
A string giving the path to the CSV file.
Use forward slashes (`/`) in file paths, even on Windows.

```r
read_csv("data/capsule_weights.csv")              # relative path
read_csv("C:/Users/Sarah/Documents/data.csv")     # absolute path
read_csv(I("col1,col2\n1,2\n3,4"))                # inline text (for demos/tests)
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> (file path or URL) | <code>connection</code> | <code>I(text)</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>col_names</code> — use first row as column names <span class="arg-types">logical | character</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
If `TRUE` (default), the first row is used as column names. If `FALSE`, column names are generated automatically (`X1`, `X2`, …). You can also supply a character vector to set names directly.

```r
read_csv("data.csv", col_names = FALSE)
read_csv("data.csv", col_names = c("id", "dose", "response"))
```

<p class="data-types"><strong>Data Types:</strong> <code>logical</code> | <code>character</code> vector · Default: <code>TRUE</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>col_types</code> — column type specification <span class="arg-types">cols() specification</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Explicitly declares the type of one or more columns.
If omitted, `read_csv()` guesses types from the first `guess_max` rows.

Common column type functions:

| Function | Type | Use when… |
|---|---|---|
| `col_character()` | text | IDs, codes, batch numbers |
| `col_double()` | decimal number | concentrations, masses, percentages |
| `col_integer()` | whole number | counts, time points in whole minutes |
| `col_logical()` | TRUE/FALSE | binary flags |
| `col_skip()` | — | ignore this column |

```r
read_csv("data.csv", col_types = cols(
  batch_id = col_character(),
  dose_mg  = col_double()
))
```

<p class="data-types"><strong>Data Types:</strong> output of <code>cols()</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>col_select</code> — select columns to read <span class="arg-types">tidy-select</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Read only a subset of columns. Accepts tidy-select syntax — column names, ranges, or helper functions like `starts_with()`.

```r
read_csv("data.csv", col_select = c(id, dose, response))
read_csv("data.csv", col_select = starts_with("conc"))
```

<p class="data-types"><strong>Data Types:</strong> tidy-select expression · Default: <code>NULL</code> (all columns)</p>
</div>
</details>

<details class="arg-item">
<summary><code>id</code> — add a column with the file path <span class="arg-types">character</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
If provided, a new column with this name is added containing the file path. Useful when reading multiple files and tracking their source.

```r
read_csv("data.csv", id = "source_file")
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> · Default: <code>NULL</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>locale</code> — locale settings for parsing <span class="arg-types">locale()</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Controls locale-specific parsing: date formats, decimal marks, encoding, and timezone. Use `locale()` to customise.

```r
read_csv("data.csv", locale = locale(decimal_mark = ",", encoding = "UTF-8"))
```

<p class="data-types"><strong>Data Types:</strong> output of <code>locale()</code> · Default: <code>default_locale()</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>na</code> — strings to interpret as missing <span class="arg-types">character vector</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
A character vector of strings that should be converted to `NA` when reading.
Default is `c("", "NA")`.
Extend this when your data uses other missing value codes.

```r
read_csv("data.csv", na = c("", "NA", "N/A", "ND", "Missing", "."))
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> vector · Default: <code>c("", "NA")</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>quoted_na</code> — treat quoted NA as missing <span class="arg-types">logical</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
If `TRUE` (default), `"NA"` inside quotes is treated as a missing value. Set to `FALSE` to keep it as the string `"NA"`.

```r
read_csv("data.csv", quoted_na = FALSE)
```

<p class="data-types"><strong>Data Types:</strong> <code>logical</code> · Default: <code>TRUE</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>quote</code> — quoting character <span class="arg-types">character</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
The single character used to quote strings. Defaults to `"\""`(double quote). Change if your file uses single quotes.

```r
read_csv("data.csv", quote = "'")
```

<p class="data-types"><strong>Data Types:</strong> single <code>character</code> · Default: <code>"\""</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>comment</code> — comment character <span class="arg-types">character</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Any text after this character on a line is ignored. Useful for files that contain comment lines starting with `#`.

```r
read_csv("data.csv", comment = "#")
```

<p class="data-types"><strong>Data Types:</strong> single <code>character</code> · Default: <code>""</code> (no comments)</p>
</div>
</details>

<details class="arg-item">
<summary><code>trim_ws</code> — trim whitespace from fields <span class="arg-types">logical</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
If `TRUE` (default), leading and trailing whitespace is stripped from each field before parsing.

```r
read_csv("data.csv", trim_ws = FALSE)
```

<p class="data-types"><strong>Data Types:</strong> <code>logical</code> · Default: <code>TRUE</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>skip</code> — rows to skip at the top <span class="arg-types">integer</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Number of lines to skip before reading the header row.
Useful when CSV files exported from instruments contain metadata rows above the column names.

```r
read_csv("instrument_export.csv", skip = 5)   # skip 5 metadata rows
```

<p class="data-types"><strong>Data Types:</strong> <code>integer</code> · Default: <code>0</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>n_max</code> — maximum number of rows to read <span class="arg-types">numeric</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Stop reading after this many data rows (not counting the header).
Useful for previewing a large file or reading only a subset.

```r
read_csv("large_dataset.csv", n_max = 100)   # first 100 rows only
```

<p class="data-types"><strong>Data Types:</strong> <code>numeric</code> · Default: <code>Inf</code> (read all rows)</p>
</div>
</details>

<details class="arg-item">
<summary><code>guess_max</code> — rows used to guess column types <span class="arg-types">integer</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Number of rows scanned to guess column types when `col_types` is not set. Increase if your file has inconsistent values in early rows.

```r
read_csv("data.csv", guess_max = 5000)
```

<p class="data-types"><strong>Data Types:</strong> <code>integer</code> · Default: <code>min(1000, n_max)</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>name_repair</code> — strategy for duplicate or invalid column names <span class="arg-types">character | function</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
How to handle column names that are duplicated or syntactically invalid. Common options: `"unique"` (default), `"minimal"`, `"universal"`, or a custom function.

```r
read_csv("data.csv", name_repair = "universal")
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> | <code>function</code> · Default: <code>"unique"</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>num_threads</code> — number of threads for parsing <span class="arg-types">integer</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Number of processing threads used when reading. Defaults to the value of `readr_threads()`, which uses all available cores.

```r
read_csv("data.csv", num_threads = 1)
```

<p class="data-types"><strong>Data Types:</strong> <code>integer</code> · Default: <code>readr_threads()</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>progress</code> — show a progress bar <span class="arg-types">logical</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
If `TRUE`, a progress bar is displayed when reading large files. Automatically suppressed in non-interactive sessions.

```r
read_csv("large_dataset.csv", progress = FALSE)
```

<p class="data-types"><strong>Data Types:</strong> <code>logical</code> · Default: <code>show_progress()</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>show_col_types</code> — suppress column type message <span class="arg-types">logical</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
By default, `read_csv()` prints a message listing each column's detected type.
Set `show_col_types = FALSE` to suppress this message in your output.

```r
data <- read_csv("data.csv", show_col_types = FALSE)
```

<p class="data-types"><strong>Data Types:</strong> <code>logical</code> · Default: <code>should_show_types()</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>skip_empty_rows</code> — skip blank rows <span class="arg-types">logical</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
If `TRUE` (default), completely empty rows are ignored. Set to `FALSE` to preserve them as rows of `NA`.

```r
read_csv("data.csv", skip_empty_rows = FALSE)
```

<p class="data-types"><strong>Data Types:</strong> <code>logical</code> · Default: <code>TRUE</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>lazy</code> — read file lazily <span class="arg-types">logical</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
If `TRUE`, the file is read lazily — only the parts that are actually accessed are loaded into memory. Useful for large files when you only need a few columns.

```r
read_csv("large_dataset.csv", lazy = TRUE)
```

<p class="data-types"><strong>Data Types:</strong> <code>logical</code> · Default: <code>should_read_lazy()</code></p>
</div>
</details>
:::

## Examples

<div class="example-section">
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="example-item">
<summary>Read a CSV file from your computer</summary>
<div class="example-body">
In practice you pass a file path:

```{webr-r}
library(readr)

data <- read_csv(
  "https://raw.githubusercontent.com/CPDSE-EDUX/SDU-CodeClub/main/debugging/dissolution_data.csv"
)

# Quick preview
head(data)
```
</div>
</details>

<details class="example-item">
<summary>Check the file structure</summary>
<div class="example-body">
See if you have to use `read_csv()` or `read_csv2()` by printing the first 3 rows of your .csv file.

```{webr-r}
read_lines("file.csv", n_max = 3)
```
</div>
</details>

<details class="example-item">
<summary>Handle missing value codes</summary>
<div class="example-body">
Clinical data exports often encode missing values as `"N/A"`, `"-"`, or a blank space. Use the optional argument `na` to unify how missing data is handled. In this example, `N/A`, `-`, and empty cells are rewritten as `NA`.

```{webr-r}
library(readr)

df_na <- read_csv(
  "data.csv",
  na = c("", "N/A", "-")
)
```
</div>
</details>

</div>
