---
title: "read_csv( )"
subtitle: "Import a CSV file as a [tibble](tibble.qmd). readcsv() and readcsv2() are from the readr package, which is part of the tidyverse."
---

<div class="source-links">
<a class="source-link" href="https://readr.tidyverse.org/reference/read_delim.html" target="_blank" rel="noopener">Original Documentation ↗</a>
<a class="source-link" href="https://rstudio.github.io/cheatsheets/html/data-import.html" target="_blank" rel="noopener">Cheatsheet ↗</a>

## Required Library
```r
install.packages("tidyverse")
library(tidyverse)
```


## Syntax

```{r, eval=FALSE}
read_csv(file)
```

```{r, eval=FALSE}
read_csv2(file)
```

### Description

::: {.desc-item}
`read_csv("file_name.csv")` reads a comma-separated values (.csv) file and returns a [tibble](tibble.qmd). Use if your file uses commas as column separators and dots for decimals.
:::
::: {.desc-item}
`read_csv2("file_name.csv")` reads a .csv file that uses semicolons as column separators and commas for decimals.
:::

::: {.callout-tip collapse="true"}
## Why 2 different functions?
English standard vs. danish standard. To find out how your `.csv`file is formatted, open it with a text editor on your computer or use the following code to print the first rows with R: `read_lines("file.csv", n_max = 3)`

```{webr-r}
read_lines("https://raw.githubusercontent.com/CPDSE-EDUX/R_documentation/refs/heads/main/assets/data/dissolution_data_raw.csv", n_max = 3)
```
:::

## Argument Overview
Required arguments must be included when using a function while optional arguments can be included on demand.

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item" open>
<summary><code>file</code> — path to the CSV file <span class="arg-types">character | connection</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
A string giving the path to the CSV file.
Use forward slashes (`/`) in file paths, even on Windows.

```r
read_csv("data/capsule_weights.csv")              # relative path
read_csv("C:/Users/Sarah/Documents/data.csv")     # absolute path
read_csv(I("col1,col2\n1,2\n3,4"))                # inline text (for demos/tests)
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> (file path or URL) | <code>connection</code> | <code>I(text)</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>col_names</code> — use first row as column names <span class="arg-types">logical | character</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
If `TRUE` (default), the first row is used as column names. If `FALSE`, column names are generated automatically (`X1`, `X2`, …). You can also supply a character vector to set names directly.

```r
read_csv("data.csv", col_names = FALSE)
read_csv("data.csv", col_names = c("id", "dose", "response"))
```

<p class="data-types"><strong>Data Types:</strong> <code>logical</code> | <code>character</code> vector · Default: <code>TRUE</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>col_select</code> — select columns to read <span class="arg-types">tidy-select</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Read only a subset of columns. Accepts tidy-select syntax — column names, ranges, or helper functions like `starts_with()`.

```r
read_csv("data.csv", col_select = c(id, dose, response))
read_csv("data.csv", col_select = starts_with("conc"))
```

<p class="data-types"><strong>Data Types:</strong> tidy-select expression · Default: <code>NULL</code> (all columns)</p>
</div>
</details>


<details class="arg-item">
<summary><code>na</code> — strings to interpret as missing <span class="arg-types">character vector</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
A character vector of strings that should be converted to `NA` when reading.
Default is `c("", "NA")`.
Extend this when your data uses other missing value codes.

```r
read_csv("data.csv", na = c("", "NA", "N/A", "ND", "Missing", "."))
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> vector · Default: <code>c("", "NA")</code></p>
</div>
</details>



<details class="arg-item">
<summary><code>comment</code> — comment character <span class="arg-types">character</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Any text after this character on a line is ignored. Useful for files that contain comment lines starting with `#`.

```r
read_csv("data.csv", comment = "#")
```

<p class="data-types"><strong>Data Types:</strong> single <code>character</code> · Default: <code>""</code> (no comments)</p>
</div>
</details>

<details class="arg-item">
<summary><code>trim_ws</code> — trim whitespace from fields <span class="arg-types">logical</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
If `TRUE` (default), leading and trailing whitespace is stripped from each field before parsing.

```r
read_csv("data.csv", trim_ws = FALSE)
```

<p class="data-types"><strong>Data Types:</strong> <code>logical</code> · Default: <code>TRUE</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>skip</code> — rows to skip at the top <span class="arg-types">integer</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Number of lines to skip before reading the header row.
Useful when CSV files exported from instruments contain metadata rows above the column names.

```r
read_csv("instrument_export.csv", skip = 5)   # skip 5 metadata rows
```

<p class="data-types"><strong>Data Types:</strong> <code>integer</code> · Default: <code>0</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>n_max</code> — maximum number of rows to read <span class="arg-types">numeric</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Stop reading after this many data rows (not counting the header).
Useful for previewing a large file or reading only a subset.

```r
read_csv("large_dataset.csv", n_max = 100)   # first 100 rows only
```

<p class="data-types"><strong>Data Types:</strong> <code>numeric</code> · Default: <code>Inf</code> (read all rows)</p>
</div>
</details>

<details class="arg-item">
<summary><code>name_repair</code> — strategy for duplicate or invalid column names <span class="arg-types">character | function</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
How to handle column names that are duplicated or syntactically invalid. Common options: `"unique"` (default), `"minimal"`, `"universal"`, or a custom function.

```r
read_csv("data.csv", name_repair = "universal")
```

<p class="data-types"><strong>Data Types:</strong> <code>character</code> | <code>function</code> · Default: <code>"unique"</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>skip_empty_rows</code> — skip blank rows <span class="arg-types">logical</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
If `TRUE` (default), completely empty rows are ignored. Set to `FALSE` to preserve them as rows of `NA`.

```r
read_csv("data.csv", skip_empty_rows = FALSE)
```

<p class="data-types"><strong>Data Types:</strong> <code>logical</code> · Default: <code>TRUE</code></p>
</div>
</details>

:::

## Examples
<a class="source-link" href="https://raw.githubusercontent.com/CPDSE-EDUX/R_documentation/refs/heads/main/assets/data/dissolution_data_raw.csv" target="_blank" rel="noopener">Example File (.csv) ↗</a>
</div>


<div class="example-section">
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="example-item">
<summary>Read a CSV file from your computer</summary>
<div class="example-body">
In practice you pass a file path:

```{webr-r}
library(tidyverse)

data <- read_csv(
  "https://raw.githubusercontent.com/CPDSE-EDUX/R_documentation/refs/heads/main/assets/data/dissolution_data_raw.csv"
)

# Quick preview
head(data)
```
</div>
</details>

<details class="example-item">
<summary>Handle missing value codes</summary>
<div class="example-body">
The [example file](https://raw.githubusercontent.com/CPDSE-EDUX/R_documentation/refs/heads/main/assets/data/dissolution_data_raw.csv) above has missing absorbance readings encoded three different ways — `"N/A"`, `"-"`, and a blank cell. Use the optional argument `na` to unify how all of them are handled, so they all become `NA`.

```{webr-r}
library(tidyverse)

df_na <- read_csv(
  "https://raw.githubusercontent.com/CPDSE-EDUX/R_documentation/refs/heads/main/assets/data/dissolution_data_raw.csv",
  skip = 2,
  na = c("", "N/A", "-")
)

df_na
```
</div>
</details>

<details class="example-item">
<summary>Skip blank rows left over from instrument exports</summary>
<div class="example-body">
The [example file](/assets/data/dissolution_data_raw.csv) also has a fully blank row between tablet 1 and tablet 2 — common after copy-pasting from Excel. By default `skip_empty_rows = TRUE` drops it automatically. Set it to `FALSE` if you need to keep it (it shows up as a row of `NA`s).

```{webr-r}
library(tidyverse)

read_csv("/assets/data/dissolution_data_raw.csv", skip = 2)
read_csv("/assets/data/dissolution_data_raw.csv", skip = 2, skip_empty_rows = FALSE)
```
</div>
</details>

<details class="example-item">
<summary>Select only the columns you need</summary>
<div class="example-body">
The dissolution data has 5 columns (`tablet`, `weight_mg`, `time_min`, `absorbance`, `dilution`). Use `col_select` to keep only the ones relevant to your analysis — here, just the tablet ID and the measured absorbance over time.

```{webr-r}
library(tidyverse)

df_subset <- read_csv(
  "https://raw.githubusercontent.com/CPDSE-EDUX/R_documentation/refs/heads/main/assets/data/dissolution_data_raw.csv",
  col_select = c(tablet, time_min, absorbance)
)

head(df_subset)
```
</div>
</details>

<details class="example-item">
<summary>Skip metadata rows from an instrument export</summary>
<div class="example-body">
The [example file](https://raw.githubusercontent.com/CPDSE-EDUX/R_documentation/refs/heads/main/assets/data/dissolution_data_raw.csv) starts with 2 lines of instrument metadata (device name, run date) before the real header row. Use `skip` to jump straight to the data — without it, `read_csv()` would try to use the device name as your column names.

```{webr-r}
library(tidyverse)

# without skip — the metadata line becomes a bogus header
read_csv("https://raw.githubusercontent.com/CPDSE-EDUX/R_documentation/refs/heads/main/assets/data/dissolution_data_raw.csv")

# with skip — the real header and data are read correctly
read_csv("https://raw.githubusercontent.com/CPDSE-EDUX/R_documentation/refs/heads/main/assets/data/dissolution_data_raw.csv", skip = 2)
```
</div>
</details>

<details class="example-item">
<summary>Preview the first rows of a large file</summary>
<div class="example-body">
Use `n_max` to read in only a handful of rows — useful for previewing a large dataset without loading the whole thing.

```{webr-r}
library(tidyverse)

df_preview <- read_csv(
  "https://raw.githubusercontent.com/CPDSE-EDUX/R_documentation/refs/heads/main/assets/data/dissolution_data_raw.csv",
  n_max = 5
)

df_preview
```
</div>
</details>

<details class="example-item">
<summary>Ignore comment lines in the file</summary>
<div class="example-body">
Some exported `.csv` files include annotation lines starting with `#`. Tell `read_csv()` to ignore everything after `#` on a line with the `comment` argument, so these annotations don't get parsed as data.

```{webr-r}
library(tidyverse)

annotated_csv <- I(
"tablet,weight_mg,time_min,absorbance,dilution
# dissolution apparatus 2, paddle speed 50 rpm
1,498,5,0.149,1
1,498,10,0.498,1
# vessel temperature held at 37C
1,498,15,0.492,2"
)

df_clean <- read_csv(annotated_csv, comment = "#")
df_clean
```
</div>
</details>

<details class="example-item">
<summary>Set your own column names</summary>
<div class="example-body">
If a file has no header row, or you want clearer column names than the original file provides, supply a character vector to `col_names`. Make sure the order matches the columns in the file, and use `skip = 1` to drop the original header row.

```{webr-r}
library(tidyverse)

df_named <- read_csv(
  "https://raw.githubusercontent.com/CPDSE-EDUX/R_documentation/refs/heads/main/assets/data/dissolution_data_raw.csv",
  col_names = c("tablet_id", "weight_mg", "time_min", "abs", "dilution_factor"),
  skip = 1
)

head(df_named)
```
</div>
</details>

</div>
