---
title: "round( )"
subtitle: "Round to a fixed number of decimal places. round() is a base R function."
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}
{{< include /_webr-options.qmd >}}

<div class="source-links">
<a class="source-link" href="https://rdrr.io/r/base/Round.html" target="_blank" rel="noopener">Original Documentation ↗</a>
</div>

## Required Library
```r
# round() is part of base R — no package installation needed
```

::: {.callout-tip collapse="true"}
## When to use round() vs. signif()
Use `round()` when a **fixed number of decimal places** is required — percentages reported to 2 d.p., pH to 2 d.p., temperatures to 1 d.p.
Use `signif()` for analytical concentrations and masses where **significant figures** reflect instrument precision.
:::

## Syntax

```{r, eval=FALSE}
round(x)
```
::: {.desc-item}
`round(x)` rounds a number (or numeric vector) to a user-specified number of decimal places. If nothing is specified, the default is 0, meaning that the result has no decimal places.

R uses **banker's rounding**: when a value is exactly halfway between two options, it rounds to the nearest *even* digit.
`round(0.5)` gives `0`, and `round(1.5)` gives `2`.
:::

::: {.callout-note}
## Only round at the point of printing
Store all intermediate values at full precision.
Call `round()` only inside `cat()` when reporting a final result.
Never call `round()` inside `mutate()` on a column that will be used in subsequent calculations.

```r
# ✓ Correct
pct_label <- ibu_mg / label_claim_mg * 100     # full precision stored
cat("Label claim:", round(pct_label, 2), "%\n") # round only when printing

# ✗ Avoid
mutate(pct_label = round(ibu_mg / label_claim_mg * 100, 2))  # rounding contaminates downstream calcs
```
:::


## Examples

<div class="example-section">
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="example-item">
<summary>Comparing <code>round()</code> to the raw value</summary>
<div class="example-body">

```{webr}
x <- 31.18765

round(x, 0)  # 0 decimals: 31
round(x, 1)  # 1 decimal: 31.2
round(x, 2)  # 2 decimals: 31.19
round(x, 3)  # 3 decimals: 31.188
round(x, 4)  # 4 decimals: 31.1876
round(x, -1) # round to tens: 30
round(x, -2) # round to hundreds: 0
```
</div>
</details>

<details class="example-item">
<summary>Positive vs. negative digits</summary>
<div class="example-body">

```{webr}
x <- 1587.346

round(x, digits =  2)    # 1587.35   — 2 decimal places
round(x, digits =  0)    # 1587      — nearest integer (default)
round(x, digits = -1)    # 1590      — nearest 10
round(x, digits = -2)    # 1600      — nearest 100
round(x, digits = -3)    # 2000      — nearest 1000
```
</div>
</details>

<details class="example-item">
<summary>Applying to a column or vector</summary>
<div class="example-body">

```{webr}
conc_values_ug_ml <- c(2.4837, 5.0124, 9.9841, 20.0152, 39.8760)

round(conc_values_ug_ml, 3)
```
</div>
</details>

<details class="example-item">
<summary><code>signif()</code> vs <code>round()</code> — which to use?</summary>
<div class="example-body">

```{webr}
x <- 0.003418

# signif — 3 sig figs regardless of decimal position
signif(x, 3)   # 0.00342  (counts from the first non-zero digit: 3, 4, 1)

# round  — fixed decimal places
round(x, 3)    # 0.003    (only 3 decimal places — loses all sig figs here!)
round(x, 5)    # 0.00342  (to get 3 sig figs you'd need to know the position)
```
</div>
</details>

</div>

## Argument Overview
<p class="arg-intro">Required arguments must be included when using a function while optional arguments can be included on demand.</p>

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>x</code> — number(s) to round <span class="arg-types">numeric | integer</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
A single number or a numeric vector.
`round()` is vectorised — it applies to every element of a vector.

<p class="data-types"><strong>Data Types:</strong> <code>numeric</code> | <code>integer</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>digits</code> — number of decimal places <span class="arg-types">integer</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Number of digits to the right of the decimal point.
Default is `0`, which rounds to the nearest integer.
Negative values round to the left of the decimal point.
```{r, eval=FALSE}
round(x, digits = 0)
```
::: {.desc-item}
`round(x, digits = 0)` adds the optional `digits` argument that controls how many decimal places to keep.
The default `0` rounds to the nearest integer; negative values round to the left of the decimal point.
:::

| `digits` | Meaning | Example: `round(1587.346, digits)` |
|---|---|---|
| `2` | 2 decimal places | `1587.35` |
| `1` | 1 decimal place | `1587.3` |
| `0` | nearest integer **(default)** | `1587` |
| `-1` | nearest 10 | `1590` |
| `-2` | nearest 100 | `1600` |

<p class="data-types"><strong>Data Types:</strong> <code>integer</code> · Default: <code>0</code></p>
</div>
</details>
:::
