---
title: "round( )"
subtitle: "Round to a fixed number of decimal places. See [original documentation](https://rdrr.io/r/base/Round.html) for reference."
---

## Syntax
Required arguments:
```{r, eval=FALSE}
round(x)
```
Advanced use — Required and optional arguments:
```{r, eval=FALSE}
round(x, digits = 0)
```

## Description

::: {.desc-item}
`round(x, digits)` rounds a number (or numeric vector) to `digits` decimal places.
Positive values of `digits` round to the right of the decimal point; negative values round to the left (e.g. `digits = -1` rounds to the nearest 10).
:::

::: {.callout-note}
## Coding convention — only round at the point of printing
Store all intermediate values at full precision.
Call `round()` only inside `cat()` when reporting a final result.
Never call `round()` inside `mutate()` on a column that will be used in subsequent calculations.

```r
# ✓ Correct
pct_label <- ibu_mg / label_claim_mg * 100     # full precision stored
cat("Label claim:", round(pct_label, 2), "%\n") # round only when printing

# ✗ Avoid
mutate(pct_label = round(ibu_mg / label_claim_mg * 100, 2))  # rounding contaminates downstream calcs
```
:::

## Examples

### Reporting percentage label claim to two decimal places

```{webr-r}
label_claim_mg <- 75.0
ibu_mg         <- 74.21   # ibuprofen content from assay (full precision)

pct_label <- ibu_mg / label_claim_mg * 100   # stored at full precision

# Two decimal places is conventional for % label claim
cat("Label claim:", round(pct_label, 2), "%\n")
```

### Reporting the acceptance value (AV) for EP 2.9.40

```{webr-r}
M    <- 100.0   # reference value (% label claim)
k_s1 <- 2.4     # EP Table 2.9.40-1, Stage 1, n = 6
L1   <- 15.0    # acceptance limit, Stage 1 (%)

X_bar_pct <- 98.4   # mean of 6 capsules
s_pct     <- 1.8    # standard deviation

AV_s1 <- abs(M - X_bar_pct) + k_s1 * s_pct  # full precision calculation

cat("Acceptance value AV:", round(AV_s1, 2), "\n")
cat("Limit L1:           ", L1, "\n")
cat("Decision:           ", ifelse(AV_s1 <= L1, "PASS ✓", "FAIL ✗"), "\n")
```

### Positive vs. negative `digits`

```{webr-r}
x <- 1587.346

round(x, digits =  2)    # 1587.35   — 2 decimal places
round(x, digits =  0)    # 1587      — nearest integer (default)
round(x, digits = -1)    # 1590      — nearest 10
round(x, digits = -2)    # 1600      — nearest 100
round(x, digits = -3)    # 2000      — nearest 1000
```

### Applying to a vector of results

```{webr-r}
pct_label_values <- c(97.832, 102.144, 99.017, 101.566, 98.253, 100.801)

# Print all values rounded to 2 decimal places
cat("% label claim:", round(pct_label_values, 2), "\n")

# Or print one per line
for (val in pct_label_values) {
  cat(" ", round(val, 2), "%\n")
}
```

### `round()` vs `signif()` — which to use?

```{webr-r}
# round() fixes decimal places — useful for percentages, pH, temperatures
round(7.438, 2)    # 7.44   — always 2 decimal places

# signif() fixes significant figures — useful for concentrations, masses
signif(7.438, 2)   # 7.4    — always 2 sig figs

# They differ most when values span orders of magnitude
values <- c(0.00412, 7.438, 1284.7)
round(values,  2)    # 0.00, 7.44, 1284.70  — many sig figs lost for small values
signif(values, 2)    # 0.0041, 7.4, 1300    — consistent precision
```

::: {.callout-tip}
## Rule of thumb
Use `round()` when a **fixed number of decimal places** is required — percentages reported to 2 d.p., pH to 2 d.p., temperatures to 1 d.p.
Use `signif()` for analytical concentrations and masses where **significant figures** reflect instrument precision.
:::

## Input Arguments

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>x</code> — number(s) to round <span class="arg-types">numeric | integer</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
A single number or a numeric vector.
`round()` is vectorised — it applies to every element of a vector.

<p class="data-types"><strong>Data Types:</strong> <code>numeric</code> | <code>integer</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>digits</code> — number of decimal places <span class="arg-types">integer</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Number of digits to the right of the decimal point.
Default is `0`, which rounds to the nearest integer.
Negative values round to the left of the decimal point.

| `digits` | Meaning | Example: `round(1587.346, digits)` |
|---|---|---|
| `2` | 2 decimal places | `1587.35` |
| `1` | 1 decimal place | `1587.3` |
| `0` | nearest integer **(default)** | `1587` |
| `-1` | nearest 10 | `1590` |
| `-2` | nearest 100 | `1600` |

<p class="data-types"><strong>Data Types:</strong> <code>integer</code> · Default: <code>0</code></p>
</div>
</details>
:::

::: {.callout-warning}
## Banker's rounding (round-half-to-even)
R uses **banker's rounding**: when a value is exactly halfway between two options, it rounds to the nearest *even* digit.
`round(0.5)` gives `0`, and `round(1.5)` gives `2`.
This is statistically unbiased and follows IEC 60559, but it may differ from the "always round up" rule you learned in school.
:::
