---
title: "signif( )"
subtitle: "Round to a given number of significant figures. See [original documentation](https://rdrr.io/r/base/Round.html) for reference."
---

## Syntax
Required arguments:
```{r, eval=FALSE}
signif(x, digits)
```
Advanced use — Required and optional arguments:
```{r, eval=FALSE}
signif(x, digits = 6)
```

## Description

::: {.desc-item}
`signif(x, digits)` rounds a number (or vector of numbers) to `digits` significant figures.
Significant figures count from the first non-zero digit, regardless of the position of the decimal point.
Use `signif()` when the *relative* precision of a measurement matters — which is the case for most analytical chemistry results.
:::

::: {.callout-note}
## Coding convention — only round at the point of printing
Store intermediate calculated values at full precision.
Call `signif()` only inside `cat()` when displaying a final result.
Rounding intermediate values introduces accumulating errors in multi-step pharmaceutical calculations.

```r
# ✓ Correct — full precision stored, rounded only when printed
C2_ug_ml <- (absorbance - std_intercept) / std_slope
cat("Concentration:", signif(C2_ug_ml, 3), "µg/mL\n")

# ✗ Avoid — rounding before using in further calculations
C2_ug_ml <- signif((absorbance - std_intercept) / std_slope, 3)
```
:::

## Examples

### Reporting an analytical concentration

```{webr-r}
# UV–Vis: calculate ibuprofen concentration from absorbance
std_intercept <- 0.003     # y-intercept of calibration curve
std_slope     <- 0.01420   # slope (AU per µg/mL)
absorbance    <- 0.445     # measured absorbance

C2_ug_ml <- (absorbance - std_intercept) / std_slope   # full precision

# Report to 3 significant figures
cat("Sample concentration:", signif(C2_ug_ml, 3), "µg/mL\n")
```

### Comparing `signif()` to the raw value

```{webr-r}
x <- 31.18765

signif(x, 1)   # 1 sig fig: 30
signif(x, 2)   # 2 sig figs: 31
signif(x, 3)   # 3 sig figs: 31.2
signif(x, 4)   # 4 sig figs: 31.19
signif(x, 5)   # 5 sig figs: 31.188
```

### Works the same for very small and very large numbers

```{webr-r}
# Sig figs count from the first non-zero digit
signif(0.0004512, 2)    # 0.00045 — first sig fig is the 4
signif(158400, 3)       # 158000
signif(0.10002, 3)      # 0.100
```

### Applying to a whole column

When used inside `cat()` with a vector, `signif()` applies to every element.
For adding a rounded column to a tibble, use `mutate()` — but only if the rounded value itself is the final deliverable (e.g., a formatted report column), not if it will be used in further calculations.

```{webr-r}
library(tibble)

conc_values_ug_ml <- c(2.4837, 5.0124, 9.9841, 20.0152, 39.8760)

# Print to 3 sig figs
cat("Concentrations (µg/mL):", signif(conc_values_ug_ml, 3), "\n")
```

### `signif()` vs `round()` — which to use?

```{webr-r}
x <- 0.003418

# signif — 3 sig figs regardless of decimal position
signif(x, 3)   # 0.00342  (counts from the first non-zero digit: 3, 4, 1)

# round  — fixed decimal places
round(x, 3)    # 0.003    (only 3 decimal places — loses all sig figs here!)
round(x, 5)    # 0.00342  (to get 3 sig figs you'd need to know the position)
```

::: {.callout-tip}
## Rule of thumb
Use `signif()` for most analytical and pharmacokinetic results where the number of significant figures is specified by the instrument or method (e.g., 3 sig figs from a UV spectrophotometer).
Use `round()` when a fixed number of decimal places is required — such as percentages always displayed to 2 decimal places.
:::

## Input Arguments

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>x</code> — number(s) to round <span class="arg-types">numeric | integer</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
A single number or a numeric vector.
`signif()` is vectorised — it applies element-wise to every value in the vector.

<p class="data-types"><strong>Data Types:</strong> <code>numeric</code> | <code>integer</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>digits</code> — number of significant figures <span class="arg-types">integer</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
The number of significant figures to keep.
Must be a positive integer.
A significant figure is any digit from the first non-zero digit onward.

| `x` | `digits` | Result | Explanation |
|---|---|---|---|
| `31.187` | `3` | `31.2` | Digits: 3, 1, 2 |
| `0.004512` | `2` | `0.0045` | Digits: 4, 5 |
| `158400` | `3` | `158000` | Digits: 1, 5, 8 |

<p class="data-types"><strong>Data Types:</strong> <code>integer</code> · Must be ≥ 1</p>
</div>
</details>
:::
