---
title: "signif( )"
subtitle: "Round to a given number of significant figures. signif() is a base R function."
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}
{{< include /_webr-options.qmd >}}

<div class="source-links">
<a class="source-link" href="https://rdrr.io/r/base/Round.html" target="_blank" rel="noopener">Original Documentation ↗</a>
</div>

## Required Library
```r
# signif() is part of base R — no package installation needed
```

::: {.callout-tip collapse="true"}
## When to use round() vs. signif()
Use `round()` when a **fixed number of decimal places** is required — percentages reported to 2 d.p., pH to 2 d.p., temperatures to 1 d.p.
Use `signif()` for analytical concentrations and masses where **significant figures** reflect instrument precision.
:::

## Syntax

```{r, eval=FALSE}
signif(x, digits)
```
::: {.desc-item}
`signif(x, digits)` rounds a number (or vector of numbers) to `digits` significant figures.
Significant figures count from the first non-zero digit, regardless of the position of the decimal point.
Use `signif()` when the *relative* precision of a measurement matters — which is the case for most analytical chemistry results.
:::

::: {.callout-note}
## Only round at the point of printing
Store intermediate calculated values at full precision.
Call `signif()` only inside `cat()` when displaying a final result.
Rounding intermediate values introduces accumulating errors in multi-step pharmaceutical calculations.

```r
# ✓ Correct — full precision stored, rounded only when printed
C2_ug_ml <- (absorbance - std_intercept) / std_slope
cat("Concentration:", signif(C2_ug_ml, 3), "µg/mL\n")

# ✗ Avoid — rounding before using in further calculations
C2_ug_ml <- signif((absorbance - std_intercept) / std_slope, 3)
```
:::


## Examples

<div class="example-section">
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="example-item">
<summary>Comparing <code>signif()</code> to the raw value</summary>
<div class="example-body">

```{webr}
x <- 31.18765

signif(x, 1)   # 1 sig fig: 30
signif(x, 2)   # 2 sig figs: 31
signif(x, 3)   # 3 sig figs: 31.2
signif(x, 4)   # 4 sig figs: 31.19
signif(x, 5)   # 5 sig figs: 31.188
```
</div>
</details>

<details class="example-item">
<summary>Applying to a whole column or vector</summary>
<div class="example-body">
When used inside `cat()` with a vector, `signif()` applies to every element.
For adding a rounded column to a tibble, use `mutate()` — but only if the rounded value itself is the final deliverable (e.g., a formatted report column), not if it will be used in further calculations.

```{webr}
conc_values_ug_ml <- c(2.4837, 5.0124, 9.9841, 20.0152, 39.8760)

signif(conc_values_ug_ml, 3)
```
</div>
</details>

<details class="example-item">
<summary><code>signif()</code> vs <code>round()</code> — which to use?</summary>
<div class="example-body">

```{webr}
x <- 0.003418

# signif — 3 sig figs regardless of decimal position
signif(x, 3)   # 0.00342  (counts from the first non-zero digit: 3, 4, 1)

# round  — fixed decimal places
round(x, 3)    # 0.003    (only 3 decimal places — loses all sig figs here!)
round(x, 5)    # 0.00342  (to get 3 sig figs you'd need to know the position)
```
</div>
</details>

</div>

## Argument Overview
<p class="arg-intro">Required arguments must be included when using a function while optional arguments can be included on demand.</p>

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>x</code> — number(s) to round <span class="arg-types">numeric | integer</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
A single number or a numeric vector.
`signif()` is vectorised — it applies element-wise to every value in the vector.

<p class="data-types"><strong>Data Types:</strong> <code>numeric</code> | <code>integer</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>digits</code> — number of significant figures <span class="arg-types">integer</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
The number of significant figures to keep.
Must be a positive integer.

```{r, eval=FALSE}
signif(x, digits = 6)
```
::: {.desc-item}
`signif(x, digits = 6)` shows the default value of `digits`.
If you omit `digits`, R keeps 6 significant figures.
In practice you should always state `digits` explicitly to match the precision of your method.
:::

A significant figure is any digit from the first non-zero digit onward.

| `x` | `digits` | Result | Explanation |
|---|---|---|---|
| `31.187` | `3` | `31.2` | Digits: 3, 1, 2 |
| `0.004512` | `2` | `0.0045` | Digits: 4, 5 |
| `158400` | `3` | `158000` | Digits: 1, 5, 8 |

<p class="data-types"><strong>Data Types:</strong> <code>integer</code> · Must be ≥ 1</p>
</div>
</details>
:::
