---
title: "slice( )"
subtitle: "Keep rows by their position. slice() is from the dplyr package, which is part of the tidyverse."
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}
{{< include /_webr-options.qmd >}}

<div class="source-links">
<a class="source-link" href="https://dplyr.tidyverse.org/reference/slice.html" target="_blank" rel="noopener">Original Documentation ↗</a>
<a class="source-link" href="https://rstudio.github.io/cheatsheets/html/data-transformation.html" target="_blank" rel="noopener">dplyr Cheatsheet ↗</a>
</div>

## Required Library
```r
install.packages("tidyverse")
library(tidyverse)
```

## Syntax

```{r, eval=FALSE}
# Pipe style
df %>% slice(row_numbers)

# Non-pipe style
slice(df, row_numbers)
```

::: {.desc-item}
`df %>% slice(2:4)` and `slice(df, 2:4)` keep rows by their position in the data,
not by a condition on the values. Negative positions drop rows instead of keeping them.
:::

```{r, eval=FALSE}
df %>% slice_max(column, n = 1)
df %>% slice_min(column, n = 1)
```

::: {.desc-item}
`slice_max()` and `slice_min()` keep the row(s) with the highest or lowest value of a column.
They are a shortcut for `arrange()` followed by `slice()`, and handle ties explicitly.
:::

::: {.callout-tip collapse="true"}
## When do you use `slice()`?
Use `slice()` when you want rows by position — for example the first three rows,
every other row, or "the row with the highest value" per group.
Use `filter()` instead when you want rows that match a logical condition on their values.
:::

## Examples

<div class="example-section">
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="example-item">
<summary>Keep rows by position</summary>
<div class="example-body">

```{webr}
library(tidyverse)

tablet_weights <- read_csv("data/tablet_weights.csv")

tablet_weights %>%
  slice(2:4)
```
</div>
</details>

<details class="example-item">
<summary>Drop rows by position</summary>
<div class="example-body">
A negative position removes that row instead of keeping it.

```{webr}
library(tidyverse)

tablet_weights <- read_csv("data/tablet_weights.csv")

tablet_weights %>%
  slice(-1)
```
</div>
</details>

<details class="example-item">
<summary>Keep the row with the highest value using <code>slice_max()</code></summary>
<div class="example-body">

```{webr}
library(tidyverse)

tablet_weights <- read_csv("data/tablet_weights.csv")

tablet_weights %>%
  slice_max(weight_mg, n = 1)
```
</div>
</details>

<details class="example-item">
<summary>Keep the first row of each group</summary>
<div class="example-body">
`slice()` respects `group_by()`, so it can pick rows within each group instead of across the whole table.

```{webr}
library(tidyverse)

tablet_weights <- read_csv("data/tablet_weights.csv")

tablet_weights %>%
  group_by(batch) %>%
  slice(1)
```
</div>
</details>

<details class="example-item">
<summary>Extract a single value with <code>pull()</code></summary>
<div class="example-body">
`slice()` always returns a data frame, even for one row.
Chain in `pull()` when you want the bare value instead of a one-row table.

```{webr}
library(tidyverse)

tablet_weights <- read_csv("data/tablet_weights.csv")

tablet_weights %>%
  slice_max(weight_mg, n = 1) %>%
  pull(weight_mg)
```
</div>
</details>

</div>

## Argument Overview
<p class="arg-intro">Required arguments must be included when using a function while optional arguments can be included on demand.</p>

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>.data</code> <span class="arg-types">data frame | tibble</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
The data frame to slice. In a pipe (`%>%`), this is passed automatically from the left-hand side.

```r
tablet_weights %>% slice(1:3)
```

<p class="data-types"><strong>Data Types:</strong> <code>data.frame</code> | <code>tibble</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>...</code> — row positions to keep <span class="arg-types">integer(s) | range</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
One or more row positions, given as numbers or a range such as `2:4`.
Use negative numbers to drop rows instead of keeping them.
Positions can't be mixed positive and negative in the same call.

```r
tablet_weights %>% slice(c(1, 3, 5))
tablet_weights %>% slice(-c(1, 2))
```

<p class="data-types"><strong>Data Types:</strong> <code>integer</code> vector or range</p>
</div>
</details>

<details class="arg-item">
<summary><code>.by</code> — group rows before slicing <span class="arg-types">column name(s)</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
Slice within each group instead of across the whole table, without needing a separate `group_by()` call.

```r
tablet_weights %>% slice(1, .by = batch)
```

<p class="data-types"><strong>Data Types:</strong> column name(s) · Default: <code>NULL</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>order_by / n</code> — used by <code>slice_max()</code> and <code>slice_min()</code> <span class="arg-types">column | integer</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
`slice_max()` and `slice_min()` take a column to rank by and an `n` for how many rows to keep.
Ties are kept by default, so more than `n` rows can be returned.

```r
tablet_weights %>% slice_max(weight_mg, n = 1)
tablet_weights %>% slice_min(weight_mg, n = 3)
```

<p class="data-types"><strong>Data Types:</strong> <code>order_by</code>: column name · <code>n</code>: <code>integer</code>, default <code>1</code></p>
</div>
</details>
:::
