---
title: "tibble( )"
subtitle: "Create a tidy data frame from column vectors. See [original documentation](https://tibble.tidyverse.org/reference/tibble.html) for reference."
---

## Syntax
Required arguments:
```{r, eval=FALSE}
tibble(col1 = values, col2 = values, ...)
```
```{r, eval=FALSE}
as_tibble(x)
```
Advanced use — Required and optional arguments:
```{r, eval=FALSE}
tibble(..., .rows = NULL, .name_repair = "check_unique")
```

## Description

::: {.desc-item}
`tibble(col1 = values, col2 = values, ...)` creates a data frame from named column vectors.
A tibble is the tidyverse version of a `data.frame` — it has the same structure but with safer defaults and cleaner printing.
:::

---

::: {.desc-item}
`tibble(col1 = values, col2 = col1 * factor)` allows later columns to reference columns defined earlier in the same call.
This is not possible with base R `data.frame()`.
:::

---

::: {.desc-item}
`as_tibble(x)` converts an existing base R `data.frame` (or matrix) into a tibble.
:::

::: {.callout-tip}
## Why tibble instead of `data.frame()`?
Tibbles never convert character columns to factors automatically, never recycle vectors of incompatible lengths silently, and print only the first 10 rows — they are designed to fail loudly and clearly, which makes bugs easier to find.
:::

## Examples

### Calibration standards table

```{webr-r}
library(tibble)

# UV–Vis calibration standards for ibuprofen assay
standards <- tibble(
  conc_ug_ml = c(0, 2.5, 5.0, 10.0, 20.0, 40.0),  # known concentrations
  absorbance  = c(0.001, 0.112, 0.224, 0.445, 0.891, 1.782)
)

standards
```

### Weighing data for content uniformity

Later columns can reference earlier ones in the same `tibble()` call.

```{webr-r}
library(tibble)

label_claim_mg    <- 75.0    # mg — labelled dose per capsule
ibu_per_mg_powder <- 0.43    # from assay

capsules <- tibble(
  capsule_id     = 1:6,
  mass_empty_mg  = c(91.2, 88.7, 90.1, 89.4, 92.0, 87.8),
  mass_filled_mg = c(265.4, 263.1, 264.8, 263.9, 266.1, 262.7),
  fill_mass_mg   = mass_filled_mg - mass_empty_mg   # references the two columns above
)

capsules
```

### Converting a base R `data.frame` to a tibble

```{webr-r}
library(tibble)

# Imagine this was loaded with read.csv() (base R)
df_base <- data.frame(
  patient_id  = c("P01", "P02", "P03"),
  weight_kg   = c(72.1, 85.4, 60.2),
  dose_mg     = c(200, 400, 200)
)

# Convert to tibble for use in a tidyverse pipeline
patients <- as_tibble(df_base)
patients
```

### Tibble printing vs. data.frame printing

```{webr-r}
library(tibble)

# A tibble shows column types and limits output to 10 rows
large_tibble <- tibble(
  id    = 1:20,
  value = rnorm(20)
)

large_tibble   # prints first 10 rows and tells you 10 more exist
```

## Input Arguments

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>...</code> — name–value pairs <span class="arg-types">name = vector</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
Each argument defines one column.
The name becomes the column name; the value must be a vector (or single value, which is recycled to the length of the longest column).

Key rules:
- All vectors must be the same length, or length 1 (recycled)
- Later arguments can reference earlier column names by name
- Column names can contain spaces if wrapped in backticks: `` `fill mass (mg)` = ... `` (not recommended — use `snake_case` with units instead)

```r
tibble(
  time_min   = c(0, 15, 30, 45, 60),   # 5 values
  dose_mg    = 200                       # recycled to 5 rows
)
```

<p class="data-types"><strong>Data Types:</strong> any vector type — <code>numeric</code>, <code>character</code>, <code>logical</code>, <code>integer</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>x</code> — object to convert (<code>as_tibble</code> only) <span class="arg-types">data.frame | matrix | list</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
The object to convert into a tibble.
Typically used to convert a `data.frame` returned by base R functions like `read.csv()` into the tidyverse format.

```r
df <- read.csv("data.csv")   # returns a data.frame
tbl <- as_tibble(df)          # convert for use with dplyr / tidyr
```

<p class="data-types"><strong>Data Types:</strong> <code>data.frame</code> | <code>matrix</code> | <code>list</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>.rows</code> — number of rows <span class="arg-types">integer</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
The number of rows the tibble should have.
Useful for creating an empty tibble with a known size:

```r
tibble(.rows = 10)   # empty tibble with 10 rows and no columns
```

<p class="data-types"><strong>Data Types:</strong> <code>integer</code></p>
</div>
</details>
:::
