---
title: "tibble( )"
subtitle: "Create a tidy data frame from column vectors. tibble() is from the tibble package, which is part of the tidyverse."
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}
{{< include /_webr-options.qmd >}}

<div class="source-links">
<a class="source-link" href="https://tibble.tidyverse.org/reference/tibble.html" target="_blank" rel="noopener">Original Documentation ↗</a>
<a class="source-link" href="https://rstudio.github.io/cheatsheets/html/tidyr.html" target="_blank" rel="noopener">tibble Cheatsheet ↗</a>
</div>

## Required Library
```r
install.packages("tidyverse")
library(tidyverse)
```

## Syntax

```{r, eval=FALSE}
tibble(col1 = values, col2 = values, ...)
```
::: {.desc-item}
`tibble(col1 = values, col2 = values, ...)` creates a data frame from named column vectors.
A tibble is the tidyverse version of a `data.frame` — it has the same structure but with safer defaults and cleaner printing.
:::

<hr>

```{r, eval=FALSE}
as_tibble(x)
```
::: {.desc-item}
`as_tibble(x)` converts an existing base R `data.frame` (or matrix) into a tibble.
:::

## Argument Overview
<p class="arg-intro">Required arguments must be included when using a function while optional arguments can be included on demand.</p>

::: {.arg-section}
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="arg-item">
<summary><code>...</code> — name–value pairs <span class="arg-types">name = vector</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
Each argument defines one column.
The name becomes the column name; the value must be a vector (or single value, which is recycled to the length of the longest column).

Key rules:
- All vectors must be the same length, or length 1 (recycled)
- Later arguments can reference earlier column names by name
- Column names can contain spaces if wrapped in backticks: `` `fill mass (mg)` = ... `` (not recommended — use `snake_case` with units instead)

```r
tibble(
  time_min   = c(0, 15, 30, 45, 60),   # 5 values
  dose_mg    = 200                       # recycled to 5 rows
)
```

<p class="data-types"><strong>Data Types:</strong> any vector type — <code>numeric</code>, <code>character</code>, <code>logical</code>, <code>integer</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>x</code> — object to convert (<code>as_tibble</code> only) <span class="arg-types">data.frame | matrix | list</span> <span class="arg-badge required">Required</span></summary>
<div class="arg-body">
The object to convert into a tibble.
Typically used to convert a `data.frame` returned by base R functions like `read.csv()` into the tidyverse format.

```r
df <- read.csv("data.csv")   # returns a data.frame
tbl <- as_tibble(df)          # convert for use with dplyr / tidyr
```

<p class="data-types"><strong>Data Types:</strong> <code>data.frame</code> | <code>matrix</code> | <code>list</code></p>
</div>
</details>

<details class="arg-item">
<summary><code>.rows</code> — number of rows <span class="arg-types">integer</span> <span class="arg-badge optional">Optional</span></summary>
<div class="arg-body">
The number of rows the tibble should have.
Useful for creating an empty tibble with a known size:

```r
tibble(.rows = 10)   # empty tibble with 10 rows and no columns
```

<p class="data-types"><strong>Data Types:</strong> <code>integer</code></p>
</div>
</details>
:::

## Examples

<div class="example-section">
<div class="collapse-toolbar"><button class="collapse-btn">Expand All</button></div>

<details class="example-item">
<summary>Create a tibble</summary>
<div class="example-body">

```{webr}
library(tidyverse)

tablet_weights <- tibble(
  tablet_id = c("T01", "T02", "T03"),
  weight_mg = c(198.4, 201.2, 199.8),
  batch = c("B2201", "B2201", "B2201")
)

tablet_weights
```
</div>
</details>

<details class="example-item">
<summary>Columns that reference earlier columns</summary>
<div class="example-body">

```{webr}
library(tidyverse)

dissolution <- tibble(
  time_min = c(5, 10, 15, 30, 45, 60),
  absorbance = c(0.12, 0.28, 0.41, 0.58, 0.67, 0.71),
  label_claim_mg = 200,
  released_mg = absorbance * 100,
  pct_dissolved = released_mg / label_claim_mg * 100
)

dissolution
```
</div>
</details>

<details class="example-item">
<summary>Repeating values</summary>
<div class="example-body">

```{webr}
library(tidyverse)

titration <- tibble(
  sample_id = 1:5,                        # insert numbers 1–5
  analyte = "Acetaminophen",              # repeat same value for all rows
  volume_titrant_ml = c(12.4, 12.6, 12.3, 12.5, 12.7)
)

titration
```
</div>
</details>

<details class="example-item">
<summary>Converting an existing data frame or list into a tibble</summary>
<div class="example-body">

```{webr}
library(tidyverse)

lims_export <- data.frame(
  sample = c("S1", "S2", "S3"),
  ph = c(6.8, 7.1, 6.9)
)
df <- as_tibble(lims_export)

df
```
</div>
</details>

<details class="example-item">
<summary>Add columns or rows</summary>
<div class="example-body">

## Original tibble
```{webr}
library(tidyverse)

# Original tibble
titration <- tibble(
  sample_id = 1:5,                        # insert numbers 1–5
  analyte = "Acetaminophen",              # repeat same value for all rows
  volume_titrant_ml = c(12.4, 12.6, 12.3, 13.2, 12.7)
)

titration
```

## Add a new row

```{webr}
library(tidyverse)

# Original tibble
titration <- tibble(
  sample_id = 1:5,                        # insert numbers 1–5
  analyte = "Acetaminophen",              # repeat same value for all rows
  volume_titrant_ml = c(12.4, 12.6, 12.3, 13.2, 12.7)
)

# Add a new row
titration <- add_row(
  titration,
  sample_id = 6, analyte = "Acetaminophen", volume_titrant_ml = 12.55
)

titration
```

## Add a new column

```{webr}
library(tidyverse)

# Original tibble
titration <- tibble(
  sample_id = 1:5,                        # insert numbers 1–5
  analyte = "Acetaminophen",              # repeat same value for all rows
  volume_titrant_ml = c(12.4, 12.6, 12.3, 13.2, 12.7)
)

# Add a new column
titration <- add_column(
  titration,
  within_spec = c(TRUE, TRUE, TRUE, FALSE, TRUE)
)

titration
```
</div>
</details>

</div>
