---
title: "R Coding Conventions"
subtitle: "A shared style guide for pharmaceutical sciences practicals"
toc: true
---
::: {.hero-banner-conventions}
# Use the Coding Conventions

You can use AI to modify existing R code according to these code conventions. Therefore, we prepared a prompt you can use. All you have to add is your R code.

[Open prompt in Claude.ai](https://claude.ai/new?q=You%20are%20an%20expert%20in%20R.%20Improve%20the%20attached%20R%20teaching%20materials%20to%20strictly%20follow%20these%20coding%20conventions%3A%20https%3A%2F%2Fraw.githubusercontent.com%2FCPDSE-EDUX%2FR_documentation%2Frefs%2Fheads%2Fmain%2FR_coding_conventions.html.%20Return%20the%20updated%20files%20and%20provide%20a%20clear%2C%20concise%20summary%20of%20all%20changes%20made%2C%20organized%20by%20category%20(e.g.%2C%20naming%2C%20indentation%2C%20comments%2C%20etc.).){.hero-btn-conventions target="_blank" rel="noopener noreferrer"}

or

[Copy a prompt to paste into any chatbot](#){.hero-btn-conventions .secondary .copy-text-btn target="_blank" rel="noopener noreferrer" data-copy-text="You are an expert in R. Improve the attached R teaching materials to strictly follow these coding conventions: https://raw.githubusercontent.com/CPDSE-EDUX/R_documentation/refs/heads/main/R_coding_conventions.html. Return the updated files and provide a clear, concise summary of all changes made, organized by category (e.g., naming, indentation, comments, etc.)."}
:::

Following these conventions makes your code readable, reproducible, and consistent across exercises.

## 1. Variable and object naming

### Use `snake_case` for all variable names

Words should be separated by underscores. Do not use dots or camelCase.

::: callout-important
This is required and used throughout all course materials.
:::

```{r, eval=FALSE}
# Do this
fill_mass_mg      <- 174.2
ibu_per_mg_powder <- 0.43
label_claim_mg    <- 75.0
X_bar_s1          <- mean(stage1$pct_label)

# Avoid this
fillMassMg <- 174.2     # camelCase
ibu.per.mg <- 0.43      # dots
lc         <- 75.0      # unclear
x1         <- mean(stage1$pct_label)
```

### Include units in variable names where relevant

Append units such as `_mg`, `_pct`, `_ug_ml`, `_min` to avoid ambiguity.

```{r, eval=FALSE}
# Do this
powder_mass_mg      <- 174.2
concentration_ug_ml <- 7.56
pct_label           <- 98.4
time_min            <- 45

# Avoid this
powder <- 174.2   # mg or g?
conc   <- 7.56    # ug/mL or mg/mL?
label  <- 98.4    # % or mg?
t      <- 45      # minutes or hours?
```

## 2. Assignment operator

### Always use `<-` for assignment, never `=`

Use `=` only for function arguments.

::: callout-important
Using `<-` consistently avoids confusion between assignment and argument naming.
:::

```{r, eval=FALSE}
# Do this
label_claim_mg <- 75.0
std_slope      <- 0.04478
data           <- readr::read_csv("file.csv")

# Avoid this
label_claim_mg = 75.0
std_slope      = 0.04478
data           = readr::read_csv("file.csv")
```

### Put spaces around `<-` and operators

```{r, eval=FALSE}
# Do this
C2           <- (absorbance - std_intercept) / std_slope
AV           <- k * S
fill_mass_mg <- mass_filled_mg - mass_empty_mg

# Avoid this
C2<-(absorbance-std_intercept)/std_slope
AV<-k*S
fill_mass_mg<-mass_filled_mg-mass_empty_mg
```

## 3. Printing results

### Use `cat()` to print labelled results

Always include a descriptive label and a unit.

```{r, eval=FALSE}
# Do this
cat("Ibuprofen content:",
		signif(ibu_per_mg_powder, 3),
		"mg Ibuprofen per mg powder\n")

cat("Mean content (X-bar):",
		signif(X_bar, 4),
		"% of label claim\n")

# Avoid this
ibu_per_mg_powder
print(ibu_per_mg_powder)
X_bar
```

### Use `view()` to inspect data frames

Use `view(data)` for interactive inspection and `head(data, 5)` for short previews.

```{r, eval=FALSE}
# Do this
view(data)
head(data, 5)

# Avoid this
print(data)
data
```

## 4. Rounding and significant figures

### Use `signif()` for final results, `round()` for fixed decimals

Keep full precision in intermediate calculations and round only when printing.

```{r, eval=FALSE}
# Store full precision
C2         <- (absorbance - std_intercept) / std_slope
ibu_per_mg <- C2 * 20 / 0.1 * 50 / 1000 / powder_mass_mg

# Round only when printing
cat("Ibuprofen per mg powder:", signif(ibu_per_mg, 3), "mg/mg\n")
cat("Label claim %:", round(pct_label, 2), "%\n")
```

### Never round inside a `mutate()` column used later

```{r, eval=FALSE}
# Do this
data %>%
	dplyr::mutate(
		fill_mass_mg = mass_filled_mg - mass_empty_mg,
		ibu_mg       = fill_mass_mg * ibu_per_mg_powder
	)

# Avoid this
data %>%
	dplyr::mutate(
		fill_mass_mg = round(mass_filled_mg - mass_empty_mg, 1),
		ibu_mg       = fill_mass_mg * ibu_per_mg_powder
	)
```

## 5. Pipes and data manipulation

### Use `%>%` with each step on its own line

One function per line, with clear indentation.

```{r, eval=FALSE}
# Do this
data <- data %>%
	dplyr::mutate(
		fill_mass_mg = mass_filled_mg - mass_empty_mg,
		ibu_mg       = fill_mass_mg * ibu_per_mg_powder,
		pct_label    = ibu_mg / label_claim_mg * 100
	)

# Avoid this
data <- dplyr::mutate(data, fill_mass_mg=mass_filled_mg-mass_empty_mg, ibu_mg=fill_mass_mg*ibu_per_mg_powder, pct_label=ibu_mg/label_claim_mg*100)
```

### Group constants at the top of a chunk

```{r, eval=FALSE}
# Do this
vol_flask1_mL  <- 50.0
vol_aliquot_mL <- 0.1
vol_flask2_mL  <- 20.0

C1 <- C2 * vol_flask2_mL / vol_aliquot_mL

# Avoid this
C1 <- C2 * 20 / 0.1
```

## 6. Plotting with ggplot2

### Use `theme_classic()` as the base theme

```{r, eval=FALSE}
ggplot2::ggplot(data, ggplot2::aes(x = time_min, y = mean_pct)) +
	ggplot2::geom_line() +
	ggplot2::geom_point() +
	ggplot2::labs(
		x = "Time (min)",
		y = "Dissolved (%)",
		title = "Dissolution profile - Ibuprofen 75 mg capsules"
	) +
	ggplot2::theme_classic()
```

### Always label axes with units in parentheses

```{r, eval=FALSE}
# Do this
ggplot2::labs(
	x = "Time (min)",
	y = "Ibuprofen released (mg)"
)

# Avoid this: no labs() call, defaults to raw variable names
```

### Put each `+` layer on its own line

```{r, eval=FALSE}
# Do this
ggplot2::ggplot(data, ggplot2::aes(x = time_min, y = mean_mg)) +
	ggplot2::geom_line() +
	ggplot2::geom_point() +
	ggplot2::geom_errorbar(
		ggplot2::aes(ymin = mean_mg - sd_mg, ymax = mean_mg + sd_mg)
	) +
	ggplot2::theme_classic()

# Avoid this
ggplot2::ggplot(data, ggplot2::aes(x=time_min,y=mean_mg))+ggplot2::geom_line()+ggplot2::geom_point()+ggplot2::theme_classic()
```

### Use `"steelblue"` as the default single-series color

For multi-group plots, map `color` in `aes(color = group)` and let ggplot choose.

```{r, eval=FALSE}
# Single series
ggplot2::geom_line(color = "steelblue")
ggplot2::geom_point(color = "steelblue")

# Multiple groups
ggplot2::ggplot(
	data,
	ggplot2::aes(x = time_min, y = mean_pct, color = formulation)
) +
	ggplot2::geom_line() +
	ggplot2::geom_point()
```

## 7. Code comments

### Comment why, not what, and always comment units

```{r, eval=FALSE}
# Useful comments
label_claim_mg <- 75.0  # mg - labelled dose per capsule
k_s1           <- 2.4   # EP 2.9.40, Table 2.9.40-1, n = 10
L1             <- 15.0  # acceptance limit, Stage 1 (%)

# Unhelpful comments
label_claim_mg <- 75.0  # set label claim to 75
k_s1           <- 2.4   # k value
L1             <- 15.0  # L1
```

### Use section separator comments in long scripts

```{r, eval=FALSE}
# --- Assay: Ibuprofen concentration from UV absorbance ---
C2 <- (absorbance - std_intercept) / std_slope
C1 <- C2 * vol_flask2_mL / vol_aliquot_mL

# --- EP 2.9.40 Stage 1 acceptance value ---
AV_s1 <- k_s1 * S_s1
```

## 8. Quick reference summary

| Topic | Convention | Level |
|------------------------|------------------------|------------------------|
| Variable names | `snake_case`, include units (`_mg`, `_pct`, `_mL`) | Required |
| Assignment | Always `<-` with spaces; `=` only inside function arguments | Required |
| Printing values | `cat("label:", signif(x, 3), "unit\n")` | Required |
| Printing tables | `view(data)` to inspect; `head(data, 5)` to preview in output | Required |
| Rounding | `signif(x, n)` or `round(x, n)` only at print-time; full precision in calculations | Required |
| Pipes | `%>%` with each step on its own indented line | Required |
| ggplot theme | Always `theme_classic()` | Required |
| Axis labels | `labs(x = "Name (unit)", y = "Name (unit)")` | Required |
| Plot layers | One `geom_*()` or modifier per line, `+` at end of line | Required |
| Default color | `"steelblue"` for single series; `aes(color = group)` for multi-series | Recommended |
| Constants | Define at top of chunk with inline unit comment | Recommended |
| Comments | Explain why and state units, not just what the code does | Required |

```{=html}
<script>
document.addEventListener('DOMContentLoaded', function () {
	var copyButtons = document.querySelectorAll('.copy-text-btn');
	if (!copyButtons.length) return;

	var copyWithFallback = function (text) {
		var helper = document.createElement('textarea');
		helper.value = text;
		helper.setAttribute('readonly', '');
		helper.style.position = 'fixed';
		helper.style.left = '-9999px';
		document.body.appendChild(helper);
		helper.focus();
		helper.select();
		var ok = false;
		try {
			ok = document.execCommand('copy');
		} catch (e) {
			ok = false;
		}
		document.body.removeChild(helper);
		return ok;
	};

	copyButtons.forEach(function (copyBtn) {
		copyBtn.addEventListener('click', function (event) {
			event.preventDefault();

			var text = copyBtn.getAttribute('data-copy-text') || '';
			if (!text) return;

			var original = copyBtn.textContent;
			var showState = function (label) {
				copyBtn.textContent = label;
				copyBtn.classList.add('copied');
				setTimeout(function () {
					copyBtn.textContent = original;
					copyBtn.classList.remove('copied');
				}, 1200);
			};

			if (navigator.clipboard && window.isSecureContext) {
				navigator.clipboard.writeText(text)
					.then(function () { showState('Copied'); })
					.catch(function () {
						var ok = copyWithFallback(text);
						showState(ok ? 'Copied' : 'Copy failed');
					});
				return;
			}

			var ok = copyWithFallback(text);
			showState(ok ? 'Copied' : 'Copy failed');
		});
	});
});
</script>
```
